
# Case Study: Heteroskedastic ANOVA {#case-ANOVA}


```{r include = FALSE}
library(tidyverse)
```
In this chapter, we present another detailed example of a simulation study, which demonstrates how to put the principles of tidy, modular simulation into practice. 
To illustrate the process of programming a simulation, we reconstruct the simulations from @brown1974SmallSampleBehavior.
We also use this case study as a recurring example in some of the following chapters.

@brown1974SmallSampleBehavior studied methods for null hypothesis testing in the studies that measure a characteristic $X$ on samples from each of several groups. They consider a population consisting of $G$ separate groups, with population means $\mu_1,...,\mu_G$ and population variances $\sigma_1^2,...,\sigma_G^2$ for the characteristic $X$.
We obtain samples of size $n_1,...,n_G$ from each of the groups, and take measurements of the characteristic for each sampled unit.
Let $x_{ig}$ denote the measurement from unit $i$ in group $g$, for $i = 1,...,n_g$ for each $j = 1,..., G$.
The analyst's goal is to use the sample data to test the hypothesis that the population means are all equal
$$
H_0: \mu_1 = \mu_2 = \cdots = \mu_G.
$$

If the population _variances_ were all equal (so $\sigma_1^2 = \sigma_2^2 = \cdots = \sigma_G^2$), we could use a conventional one-way analysis of variance (ANOVA) to test.
However, one-way ANOVA might not work well if the variances are not equal.
The question is then what are best practices for testing the null of equal group means, allowing for the possibility that variances could differ across groups (a form of heteroskedasticity).

To tackle this question, Brown and Forsythe evaluated two different hypothesis testing procedures, developed by @james1951ComparisonSeveralGroups and @welch1951ComparisonSeveralMean, which avoid the assumption that all groups have equal equality of variances.
Brown and Forsythe also evaluated the conventional one-way ANOVA F-test as a benchmark, even though this procedure maintains the assumption of equal variances.
They also proposed and evaluated a new procedure of their own devising.[^typical-stats-methods]
Overall, the simulation involves comparing the performance of these different hypothesis testing procedures (the methods) under a range of conditions (different data generating processes) with different sample sizes and different degrees of heteroskedasticity.

[^typical-stats-methods]: This latter piece makes Brown and Forsythe's study a prototypical example of a statistical methodology paper: find some problem that current procedures do not perfectly solve, invent something to do a better job, and then do simulations and/or math to build a case that the new procedure is better.

When evaluating hypothesis testing procedures, there are two main performance metrics of interest: type-I error rate and power.
The type-I error rate is the rate at which a test rejects the null hypothesis when the null hypothesis is actually true.
To apply a hypothesis testing procedure, one has to specify a desired, or nominal, type-I error rate, often denoted as the $\alpha$-level. 
For a specified $\alpha$, a valid or well-calibrated test should have an actual type-I error rate less than or equal to the nominal level, and ideally should be very close to nominal. 
<!-- JEP: Do you have a source for the terminology of a _valid_ test? I'm kind of wary of calling it the validity of a test because validity is already such an overloaded term. In some of my work, I've written about type-I error as measuring whether the test is appropriately _calibrated_, which lines up with the idea of calibrating physical measurement instruments like scales or temperature gauges. -->
Power is how often a test correctly rejects the null when it is indeed false.
It is a measure of how _powerful_ (or sensitive) a method is.

Brown and Forsythe estimated error rates and power for nominal $\alpha$-levels of 1%, 5%, and 10%.
Table 1 of their paper reports the simulation results for type-I error (labeled as "size").
They looked at ten different scenarios:

```{r BF-Scenarios, echo=FALSE, message = FALSE}
library(kableExtra)

table_1_scenarios <- tribble(
  ~Groups, ~`Sample Sizes`,           ~`Standard Deviations`,
  "4",     "4,4,4,4",                 "1,1,1,1",
  "4",     "4,8,10,12",               "1,2,2,3",
  "4",     "11,11,11,11",             "1,1,1,1",
  "4",     "11,16,16,21",             "1,2,2,3",
  "6",     "4,4,4,4,4,4",             "1,1,1,1,1,1",
  "6",     "4,6,6,8,10,12",           "1,2,2,3",
  "6",     "6,6,6,6,6,6",             "1,1,1,1,1,1",
  "6",     "11,11,11,11,11,11",       "1,1,1,1,1,1",
  "6",     "16,16,16,16,16,16",       "1,2,2,3",
  "6",     "21,21,21,21,21,21",       "1,2,2,3",
  "10",    "20,20,20,20,20,20,20,20,20,20", "1,1,1.5,1.5,2,2,2.5,2.5,3,3"
) %>%
  mutate(
    Scenario = LETTERS[1:11]
  )

table_1_scenarios %>%
  relocate(Scenario) %>%
  kbl( 
    caption = "Simulation scenarios explored by Brown and Forsythe (1974)"
  ) %>%
  kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped","compact","hover")
  ) 
```

We also provide some of the numbers they reported in their Table 1 for these scenarios on Table \@ref(tab:BF-table1).

```{r BF-table1, echo=FALSE}
library(tibble)

table_5_percent <- tribble(
  ~Groups, ~`F`, ~`F*`, ~`W`, ~`J`,
  "4",     4.9,     3.4,      4.5,     7.9,
  "4",     6.7,     4.1,      4.7,     8.1,
  "4",     5.1,     4.8,      5.7,     6.7,
  "4",     3.0,     5.7,      4.9,     5.6,
  "6",     14.4,    6.2,      6.5,     7.7,
  "6",     5.1,     5.7,      5.0,     5.5,
  "6",     6.3,     5.7,      5.0,     5.4,
  "6",     4.9,     5.1,      5.0,     5.3,
  "6",     10.8,    6.2,      5.5,     5.8,
  "6",     4.0,     6.5,      5.4,     5.6,
  "10",    4.9,     3.4,      6.1,     9.5
)
#nrow( table_5_percent )
table_1_percent <- tribble(
  ~Groups, ~`F`, ~`F*`, ~`W`, ~`J`,
  "4",     0.9,     0.5,      0.8,     2.4,
  "4",     1.7,     0.7,      0.8,     2.7,
  "4",     1.1,     1.0,      1.6,     2.1,
  "4",     0.6,     1.4,      0.9,     1.3,
  "6",     5.6,     1.8,      2.0,     2.9,
  "6",     1.0,     1.5,      0.9,     1.1,
  "6",     1.8,     1.5,      1.1,     1.3,
  "6",     1.1,     1.0,      1.0,     1.2,
  "6",     3.9,     1.8,      1.2,     1.3,
  "6",     1.0,     1.8,      1.1,     1.1,
  "10",    1.0,     0.4,      1.4,     3.8
)


table_10_percent <- tribble(
  ~Groups, ~`F`, ~`F*`, ~`W`, ~`J`,
  "4",     10.2,     7.8,       9.6,      13.3,
  "4",     12.0,     8.9,       10.3,     13.8,
  "4",     9.9,      9.5,       10.8,     12.1,
  "4",     5.9,      10.3,      9.8,      10.8,
  "6",     21.9,     11.0,      11.3,     12.9,
  "6",     10.1,     9.8,       10.0,     10.6,
  "6",     11.4,     10.7,      10.1,     10.6,
  "6",     10.3,     10.3,      10.2,     10.5,
  "6",     17.3,     11.1,      10.5,     10.9,
  "6",     7.3,      11.5,      10.6,     10.9,
  "10",    9.6,      7.3,       11.4,     14.7
)

table_1 <- bind_rows(
  `10%` = table_10_percent,
  `5%` = table_5_percent,
  `1%` = table_1_percent,
  .id = "level"
)

table_1 <- 
  table_1 %>%
  mutate(
    Scenario = rep(LETTERS[1:11], 3),
    level = factor(level, levels = c("10%", "5%", "1%"))
  ) %>%
  pivot_wider(
    names_from = level, 
    values_from = c(`F`, `F*`, W, J)
  ) %>%
  relocate(Scenario)

table_1 %>%
  kbl( 
    caption='Portion of "Table 1," reproduced from Brown and Forsythe',
    col.names = c("Scenario", "Groups",rep(c("10%","5%","1%"), times = 4))
  ) %>%
  kable_styling(
    full_width = FALSE, 
    bootstrap_options = c("striped","compact","hover"),
  ) %>%
  add_header_above(c(" " = 2, "ANOVA F test" = 3, "Brown & Forsythe F* test" = 3,"Welch's test" = 3, "James' test" = 3))
```

Table \@ref(tab:BF-table1) reports results on power; it is desirable to have higher power to reject null hypotheses that are false, so higher rates are better here.

To replicate this simulation, we will first write functions to generate data and evaluate the hypothesis testing procedures in a specific scenario with a specific set of core parameters (e.g., sample sizes, number of groups, and so forth). We will then scale up to do a range of scenarios where we vary the parameters of the data-generating model


## The data-generating model

In the heteroskedastic one-way ANOVA simulation, there are three sets of parameter values: population means, population variances, and sample sizes.
Rather than attempting to write a general data-generating function immediately, it is often easier to write code for a specific case first and then use that code as a starting point for developing a function.
For example, say that we have four groups with means of 1, 2, 5, 6; variances of 3, 2, 5, 1; and sample sizes of 3, 6, 2, 4:

```{r}
mu <- c(1, 2, 5, 6)
sigma_sq <- c(3, 2, 5, 1)
sample_size <- c(3, 6, 2, 4)
```

Following @brown1974SmallSampleBehavior, we will assume that the measurements are normally distributed within each sub-group of the population. The following code generates a vector of group id's and a vector of simulated measurements:

```{r}
N <- sum(sample_size) # total sample size
G <- length(sample_size) # number of groups

# group id factor
group <- factor(rep(1:G, times = sample_size))

# mean for each unit of the sample
mu_long <- rep(mu, times = sample_size) 

# sd for each unit of the sample
sigma_long <- rep(sqrt(sigma_sq), times = sample_size) 

# See what we have?
tibble( group = group, mu = mu_long, sigma = sigma_long )

# Now make our data
x <- rnorm(N, mean = mu_long, sd = sigma_long)
dat <- tibble(group = group, x = x)
dat
```

We have made a small dataset of group membership and outcome.[^many-ways]
We have followed the strategy of first constructing a dataset with parameters for each observation in each group, making heavy use of base R's `rep()` function to repeat values in a list.
We then called `rnorm()` to generate `N` observations in all. 
This works because `rnorm()` is _vectorized_; if you give it a vector or vectors of parameter values, it will generate each subsequent observation according to the next entry in the vector. As a result, the first `x` value is simulated from a normal distribution with mean `mu_long[1]` and standard deviation `sd_long[1]`, the second `x` is simulated using `mu_long[2]` and `sd_long[2]`, and so on.

[^many-ways]: As usual, there are many different and legitimate ways of doing this in R.
For instance, instead of using `rep` to do it all at once, we could generate observations for each group separately, then stack the observations into a dataset.
Do not worry about trying to writing code the "best" way---especially when you are initially putting a simulation together.
We strongly believe in adage that if you can do it at all, then you should feel good about yourself.

### Now make a function

Because we will need to generate datasets over and over, we wrap our code in a function.
The inputs to the function will be the parameters of the model that we specified at the very beginning: the set of population means `mu`, the population variances `sigma_sq`, and sample sizes `sample_size`. We make these quantities arguments of the data-generating function so that we can make datasets of different sizes and shapes:

```{r generate_data_block}
generate_data <- function(mu, sigma_sq, sample_size) {

  N <- sum(sample_size)
  G <- length(sample_size)

  group <- factor(rep(1:G, times = sample_size))
  mu_long <- rep(mu, times = sample_size)
  sigma_long <- rep(sqrt(sigma_sq), times = sample_size)

  x <- rnorm(N, mean = mu_long, sd = sigma_long)
  sim_data <- tibble(group = group, x = x)

  return(sim_data)
}
```

The above code is simply the code we built previously, all bundled up.
We developed the function by first writing code to make the data-generating process to work once, the way we want, and only then turning the final code into a function for later reuse.

Once we have turned the code into a function, we can call to get a new set of simulated data.
For example, to generate a dataset with the same parameters as before, we can do:

```{r}
sim_data <- generate_data(
  mu = mu, 
  sigma_sq = sigma_sq,
  sample_size = sample_size
)

sim_data
```

To generate one with population means of zero in all the groups, but the same group variances and sample sizes as before, we can do:

```{r}
sim_data_null <- generate_data(
  mu = c( 0, 0, 0, 0 ),
  sigma_sq = sigma_sq, 
  sample_size = sample_size
)

sim_data
```

Following the principles of tidy, modular simulation, we have written a function that returns a rectangular dataset for further analysis. 
Also note that the dataset returned by `generate_data()` only includes the variables `group` and `x`, but not `mu_long` or `sd_long`. 
This is by design. 
Including `mu_long` or `sd_long` would amount to making the population parameters available for use in the data analysis procedures, which is not something that happens when analyzing real data.

### Cautious coding

In the above, we built some sample code, and then bundled it into a function by literally cutting and pasting the initial work we did into a function skeleton.
In the process, we shifted from having variables in our workspace with different names to using those variable names as parameters in our function call.

Developing code in this way is not without hazards.
In particular, after finishing making our function, our workspace has a variable `mu` in it and our function also has a parameter named `mu`.
Inside the function, R will use the parameter `mu` first, but this is potentially confusing.
As are, potentially, lines such as `mu = mu`, which means "set the function's parameter called `mu` to the variable called `mu`."
These are different things (with the same name).

One way to check your code, once a function is built, is to comment out the initial code (or delete it), restart R or at least clear out the workspace, and then re-run the code that uses the function.
If things still work, then you should be somewhat confident you successfully bundled your code into the function.

You can also, once you bundle your code, do a search and replace to change variable names in your function to something more generic, to make the separation more clear.

## The hypothesis testing procedures

Brown and Forsythe considered four different hypothesis testing procedures for heteroskedastic ANOVA, but we will focus on just two of the tests for now.
We start with the conventional one-way ANOVA while mistakenly assuming homoskedasticity. 
R's `oneway.test` function will calculate this test automatically:

```{r}
sim_data <- generate_data(
  mu = mu, 
  sigma_sq = sigma_sq,
  sample_size = sample_size
)

anova_F <- oneway.test(x ~ group, data = sim_data, var.equal = TRUE)
anova_F
```
We can use the same function to calculate Welch's test by setting `var.equal = FALSE`:
```{r}
Welch_F <- oneway.test(x ~ group, data = sim_data, var.equal = FALSE)
Welch_F
```
The main results we need here are the $p$-values of the tests, which will let us assess Type-I error and power for a given nominal $\alpha$-level. The following function takes simulated data as input and returns as output the $p$-values from the one-way ANOVA test and Welch test:

```{r}
ANOVA_Welch_F <- function(data) {
  anova_F <- oneway.test(x ~ group, data = data, var.equal = TRUE)
  Welch_F <- oneway.test(x ~ group, data = data, var.equal = FALSE)
  
  result <- tibble(
    ANOVA = anova_F$p.value,
    Welch = Welch_F$p.value
  )
  
  return(result)
}

ANOVA_Welch_F(sim_data)
```
Following our tidy, modular simulation principles, this function returns a small dataset with the p-values from both tests. 
Eventually, we might want to use this function on some real data.
Our estimation function does not care if the data are simulated or not; we call the input `data` rather than `sim_data` to reflect this.

As an alternative to this function, we could instead write code to implement the ANOVA and Welch tests ourselves.
This has some potential advantages, such as avoiding any extraneous calculations that `oneway.test` does, which take time and slow down our simulation.
However, there are also drawbacks to doing so, including that writing our own code takes _our_ time and opens up the possibility of errors in our code.
For further discussion of the trade-offs, see Chapter \@ref(optimize-code), where we do implement these tests by hand and see what kind of speed-ups we can obtain.

## Running the simulation

We now have functions that implement steps 2 and 3 of the simulation.
Given some parameters, `generate_data` produces a simulated dataset and `ANOVA_F_aov` and `Welch_F` use the simulated data to calculate $p$-values two different ways.
We now want to know which way is better, and how.
To answer this question, we next need to repeat this chain of calculations a bunch of times.

We first make a function that puts our chain together in a single method:
```{r}
one_run = function( mu, sigma_sq, sample_size ) {
  sim_data <- generate_data(
    mu = mu, 
    sigma_sq = sigma_sq,
    sample_size = sample_size
  )
  ANOVA_Welch_F(sim_data)
}

one_run( mu = mu, sigma_sq = sigma_sq, sample_size = sample_size )
```
This function implements a single simulation trial by generating artificial data and then analyzing the data, ending with a nice dataframe or tibble that has results for the single run.

We next call `one_run()` over and over; see \@ref(repeating-oneself) for some discussion of options.
The following uses `map_df` to run `one_run()` 4 times and then stack the results into a single data frame:

```{r}
sim_data <- map_df(
  1:4, 
  ~ one_run(
      mu = mu, 
      sigma_sq = sigma_sq, 
      sample_size = sample_size
    )
)
sim_data
```

Voila! We have simulated $p$-values!

## Summarizing Test Performance

We now have all the pieces in place to reproduce the results from Brown and Forsythe (1974).
We first focus on calculating the actual type-I error rate of these tests---that is, the proportion of the time that they reject the null hypothesis of equal means when that null is actually true---for an $\alpha$-level of .05.
We therefore need to simulate data according to process where the population means are indeed all equal. Arbitrarily, we start with $G = 4$ groups and set all of the means equal to zero:

```{r}
mu <- rep(0, 4)
```

In the fifth row of Table 1, Brown and Forsythe examine performance for the following parameter values for sample size and population variance:

```{r}
sample_size <- c(4, 8, 10, 12)
sigma_sq <- c(3, 2, 2, 1)^2
```

With these parameter values, we can use `map_dfr` to simulate 10,000 $p$-values:

```{r welch_sim_1, cache=TRUE}
p_vals <- map_dfr(1:10000, 
  ~ one_run(
      mu = mu,
      sigma_sq = sigma_sq,
      sample_size = sample_size
    ) 
)

p_vals
```

We next use our replications to calculate the rejection rates.
The rule is that the null is rejected if the $p$-value is less than $\alpha$. To get the rejection rate, calculate the proportion of replications where the null is rejected.
```{r}
sum(p_vals$ANOVA < 0.05) / 10000
```
This is equivalent to taking the mean of the logical conditions:

```{r}
mean(p_vals$ANOVA < 0.05)
```

We get a rejection rate that is much larger than $\alpha = .05$.
We have learned that the ANOVA F-test does not adequately control Type-I error under this set of conditions.

```{r}
mean(p_vals$Welch < 0.05)
```

The Welch test does much better, although it appears to be a little bit in excess of 0.05.

Note that these two numbers are quite close (though not quite identical) to the corresponding entries in Table 1 of Brown and Forsythe (1974). The difference is due to the fact that both Table 1 and are results are actually _estimated_ rejection rates, because we have not actually simulated an infinite number of replications. The estimation error arising from using a finite number of replications is called _simulation error_ (or _Monte Carlo error_).
Later on in Chapter \@ref(performance-criteria), we will look more at how to estimate and control the Monte Carlo simulation error in performance measures.

## Exercises {#exAnovaExercises}

The following exercises involve exploring and tweaking the above simulation code we have developed to replicate the results of Brown and Forsythe (1974).

1. Table 1 from Brown and Forsythe reported rejection rates for $\alpha = .01$ and $\alpha = .10$ in addition to $\alpha = .05$. Calculate the rejection rates of the ANOVA F and Welch tests for all three $\alpha$-levels.

2. Try simulating the Type-I error rates for the parameter values in the first two rows of Table 1 of the original paper. Use 10,000 replications. How do your results compare to the results reported in Table 1?

3. Try simulating the **power levels** for a couple of sets of parameter values from Table 2. Use 10,000 replications. How do your results compare to the results reported in the Table?

4. Instead of making `ANOVA_Welch_F` return a single row with the columns for the $p$-values, one could instead return a dataset with one row for each test. The "long" approach is often nicer when evaluating more than two methods, or when each method returns not just a $p$-value but other quantities of interest. In this example, we might also want to store the $F$ statistic The resulting dataset would then look like the following:

    ```{r, include=FALSE}
    ANOVA_Welch_F_long <- function(data) {
      anova_F <- oneway.test(x ~ group, data = data, var.equal = TRUE)
      Welch_F <- oneway.test(x ~ group, data = data, var.equal = FALSE)
      
      tibble(
        method = c( "ANOVA", "Welch" ),
        Fstat = c( anova_F$statistic, Welch_F$statistic),
        pvalue = c( anova_F$p.value, Welch_F$p.value )
      )
      
    }
    ```

    ```{r, echo=TRUE}
    ANOVA_Welch_F_long(sim_data)
    ```

    Modify `ANOVA_Welch_F()` to do this, update your simulation code, and then use `group_by()` plus `summarise()` to calculate rejection rates of both tests.

5. The `onewaytests` package in R includes functions for calculating Brown and Forsythe's $F^*$ test and James' test for differences in population means. Modify the data analysis function `ANOVA_Welch_F` (or `ANOVA_Welch_F_long`) to also include results from these hypothesis tests. Re-run the simulation to estimate the type-I error rate of all four tests under Scenarios A and B of Table \@ref(tab:BF-Scenarios)
