---
output:
  pdf_document: default
  html_document: default
---

```{r include = FALSE}
library(tidyverse)
options(list(dplyr.summarise.inform = FALSE))

### Code from the prior chapter
source( "case_study_code/clustered_data_simulation.R" )

dat <- gen_dat_model( n=5, J=3, p=0.5, 
                      gamma_0=0, gamma_1=0.2, gamma_2=0.2,
                      sigma2_u = 0.4, sigma2_e = 1,
                      alpha = 0.5 )

```


# Running the Simulation

In the prior two chapters we saw how to write functions that generate data according to a specified model (and parameters) and functions that implement estimation procedures on simulated data.
We next put those two together and repeat a bunch of times to obtain a lot of results such as point estimates, estimated standard errors and/or confidence intervals.

We use two primary ways of doing this in this textbook.
The first is to write a function that does a single step of a simulation, and then use the `map()` function to run that single step multiple times.

For our Cluster RCT case study, for example, we would write the following that takes our simulation parameters and runs a single trial of our simulation:
```{r}
one_run <- function( n_bar = 30, J=20, 
                     gamma_1 = 0.3, gamma_2 = 0.5,
                     sigma2_u = 0.20, sigma2_e = 0.80,
                     alpha = 0.75 ) {
  
  dat <- gen_dat_model( n_bar = n_bar, J=J, 
                        gamma_1 = gamma_1, gamma_2 = gamma_2,
                        sigma2_u = sigma2_u, sigma2_e = sigma2_e,
                        alpha = alpha )
  MLM = analysis_MLM( dat )
  LR = analysis_OLS( dat )
  Agg = analysis_agg( dat )
  
  bind_rows( MLM = MLM, LR = LR, Agg = Agg,
             .id = "method" )
}
```

We have added a bunch of defaults to our function, so we can easily run it without remembering all the things we can change.

When we call it, we get a nice table of results that we can evaluate:

```{r, messages=FALSE}
one_run( n_bar = 30, J = 20, alpha=0.5 )
```
The results for each method is a single line.
We record estimated impact, estimated standard error, and a nominal $p$-value.
Note how the `bind_rows()` method can take naming on the fly, and give us a column of `method`, which will be very useful for keeping track of what estimated what.
We intentionally wrap up our results with a data frame to make later processing of data with the tidyverse package much easier.

We then use the `map()` function to run this function multiple times:

```{r secret_run_cluster_rand_sim, include=FALSE}
R <- 1000
ATE <- 0.30

if ( !file.exists("results/cluster_RCT_simulation.rds") ) {
  tictoc::tic()  # Start the clock!
  set.seed( 40404 )
  runs <- 
    map_df( 1:R, ~one_run( n_bar = 30, J=20, gamma_1 = ATE ) ) 
  
  tictoc::toc()
  
  saveRDS( runs, file = "results/cluster_RCT_simulation.rds" ) 
} else {
  runs = readRDS( file = "results/cluster_RCT_simulation.rds" ) 
}
```


```{r cluster_rand_sim, eval=FALSE}
set.seed( 40404 )
R = 1000
ATE = 0.30
runs <- 
  map_df( 1:R, ~one_run( n_bar = 30, J=20, gamma_1 = ATE ),
          .id="runID" ) 

saveRDS( runs, file = "results/cluster_RCT_simulation.rds" )
```

What the `map()` function is doing is first making a list from 1 to R, and then for each element in that list, it is calling `one_run()` with the parameters `n_bar = 30, J=20`.
The `~` is a shorthand way of writing a function that takes one argument, and then calls `one_run()` with that argument; the argument is the iteration number (1, 2, 3, ..., R), but we are ignoring it.  The `.id = "runID"` argument is a way of keeping track of which iteration number produced which result.
The `_df` at the end of `map_df()` is a way of telling `map()` to take the results of each iteration and bind them together into a single data frame.

Once our simulation is complete, we save our results to a file for future use; this speeds up our lives since we will not have to constantly re-run our simulation each time we want to explore the results.

We have arrived!  We now have the individual results of all our methods applied to each of 1000 generated datasets.
The next step is to evaluate how well the estimators did.
Regarding our point estimate, for example, we have these primary questions:

 - Is it biased? (bias)
 - Is it precise? (standard error)
 - Does it predict well? (RMSE)
 
In the next chapter, we systematically go through answering these questions for our initial scenario.


## Writing simulations quick with the simhelpers package

The `map` approach is a bit strange, with building a secret function on the fly with `~`, and also having the copy over all the parameters we pass from `one_run()` to `gen_dat_model()`.
The `simhelpers` package provides a shortcut that makes this step easier.

To do it, we first need to write a single estimation procedure function that puts all of our estimators together:
  
```{r, message = FALSE}
analyze_data = function( dat ) {
  MLM = analysis_MLM( dat )
  LR = analysis_OLS( dat )
  Agg = analysis_agg( dat )
  
  bind_rows( MLM = MLM, LR = LR, Agg = Agg,
             .id = "method" )
}

```

This is simply the `one_run()` method from above, but without the data generating part.
When we pass a dataset to it, we get a nice table of results that we can evaluate, as we did before.

```{r, messages=FALSE}
dat = gen_dat_model( n=30, J = 20, gamma_1 = 0.30 )
analyze_data( dat )
```

We can now use `simhelpers` to write us a new function for the entire simulation:
```{r, messages=FALSE}
library(simhelpers)
sim_function <- bundle_sim( gen_dat_model, analyze_data )
```

We can then use it as so:
```{r, messages=FALSE}
sim_function( 2, n_bar = 30, J = 20, gamma_1 = ATE )
```

The `bundle_sim()` command takes our DGP function and our estimation procedures function and gives us back a function, which we have called `sim_function`, that will run a simulation using whatever parameters we give it.
The `bundle_sim()` command examines `gen_dat_model` function, figures out what parameters it needs, and makes sure that the newly created function is able to take those parameters from the user.

To use it for our simulation, we would then write
```{r, eval=FALSE}
rns <- sim_function( R, n_bar = 30, J = 20, gamma_1 = ATE )
saveRDS( runs, file = "results/cluster_RCT_simulation.rds" )
```

This is a bit more elegant than the `map()` approach, and is especially useful when we have a lot of parameters to pass around.


## Adding Checks and Balances

In the extensions of the prior DGP chapter, we discussed indexing our DGP by the ICC instead of the two variance components.
We can do this, and also translate some of the more obscure model parameters to easier to interpret parameters from within our simulation driver as follows:

```{r revised_CRT, eval=FALSE}
one_run <- function( n_bar = 30, J=20, 
                     ATE = 0.3, size_coef = 0.5,
                     ICC = 0.4,
                     alpha = 0.75 ) {
    stopifnot( ICC >= 0 && ICC < 1 )

  dat <- gen_dat_model( n_bar = n_bar, J=J, 
                        gamma_1 = ATE, gamma_2 = size_coef,
                        sigma2_u = ICC, sigma2_e = 1-ICC,
                        alpha = alpha )
  MLM = analysis_MLM( dat )
  LR = analysis_OLS( dat )
  Agg = analysis_agg( dat )
  
  bind_rows( MLM = MLM, LR = LR, Agg = Agg,
             .id = "method" )
}
```

Note the `stopifnot`: it is wise to ensure our parameter transforms are all reasonable, so we do not get unexplained errors or strange results later on.
It is best if your code fails as soon as possible!  Otherwise debugging can be quite hard.

In our modified `one_run()` we are transforming our ICC parameter into specific other parameters that are used in our actual model to maintain our effect size interpretation of our simulation.
We have not even modified our `gen_dat_model()` DGP method: we are just specifying the constellation of parameters as a function of the parameters we want to directly control in the simulation.

Controlling how we use the foundational elements such as our data generating code is a key tool for making the higher level simulations sensible and more easily interpretable.
Here we have put our entire simulation into effect size units, and are now providing "knobs" to the simulation that are directly interpretable.


## Exercises


1. In the prior chapter's exercises, you made a new `BF_F` function for the Welch simulation. Now incorporate the `BF_F` function into the `one_run()` function, and use your revised function to generate simulation results for this additional estimator.




