```{r, include=FALSE}
library( blkvar )
```

# More on functions {#functions}

This chapter is not about simulation, but does have a few tips and tricks regarding coding that are worth attending to.

## Default arguments for functions {#default_arguments}

To generate easy-to-use, but easy to configure code, use default arguments.
For example,

```{r}
my_function = function( a = 10, b = 20 ) {
     100 * a + b
}

my_function()
my_function( 5 )
my_function( b = 5 )
my_function( b = 5, a = 1 )
```

We can call it when we don't know what the arguments are, but then when we know more about the function, we can specify things of interest.
Lots of R commands work exactly this way, and for good reason.

Especially for code to generate random datasets, default arguments can be a lifesaver as you can call the method before you know exactly what everything means.

For example, consider the `blkvar` package that has some code to generate blocked randomized datasets.
We might locate a promising method, and type it in:

```{r, error=TRUE}
library( blkvar )
generate_blocked_data()
```

That didn't work, but let's provide some block sizes and see what happens:

```{r}
generate_blocked_data( c( 3, 2 ) )
```

Nice!  We see that we have a block ID and the control and treatment potential outcomes.  We also don't see a random assignment variable, so that tells us we probably need some other methods as well.
But we can play with this as it stands right away.

Next we can see that there are many things we might tune:
```{r}
args( generate_blocked_data )
```

The documentation will tell us more, but if we just need some sample data, we can quickly assess our method before having to do much reading and understanding.
Only once we have identified what we need do we really need to turn to the documentation itself.

