# Error trapping and other headaches

If you have an advanced estimator, or are relying on some package, it is quite possible that every so often your estimate will trigger an error, or give you a NA result, or something similarly bad.
More innocuous, you might have estimators that can generate warnings; if you run 10,000 trials, that can add up to a lot of warnings which can be overwhelming to sort through.

In this section, we talk about some ways to make your simulations safe and robust.

## Safe code


## Saving files and results

Always save your simulation results to a file.
They are painful and time consuming to run, and you will invariably want to analyze them in a variety of different ways, once you have looked at your preliminary analysis.
We advocate saving your simulation as soom as it is complete.
But there are some ways to do better than that, such as saving as you go.

### Saving simulations as you go ###

If you are not sure you have time to run your entire simulation, or you think your computer might crash half way through, or something similar, you can save each chunk you run as you go, in its own file.  You then stack those files at the end to get your final results.
With clever design, you can even then selectively delete files to rerun only parts of your larger simulation---but be sure to rerun everything from scratch before you run off and publish your results, to avoid embarrassing errors.

Here, for example, is a script from a research project examining how one might use post-stratification to improve the precision of an IV estimate.
This is the script that runs the simulation.
Note the sourcing of other scripts that have all the relevant functions; these are not important here.
Due to modular programming, we can see what this script does, even without those detail.

```{r IV_simulation_driver, eval=FALSE}
source( "pack_simulation_functions.R" )

if ( !file.exists("results/frags" ) ) {
    dir.create("results/frags")
}

# Number of simulation replicates per scenario
R = 1000

# Do simulation breaking up R into this many chunks
M_CHUNK = 10

###### Set up the multifactor simulation #######

# chunkNo is a hack to make a bunch of smaller chunks for doing parallel more
# efficiently.
factors = expand_grid( chunkNo = 1:M_CHUNK,
                       N = c( 500, 1000, 2000 ),
                       pi_c = c( 0.05, 0.075, 0.10 ),
                       nt_shift = c( -1, 0, 1 ),
                       pred_comp = c( "yes", "no" ),
                       pred_Y = c( "yes", "no" ),
                       het_tx = c( "yes", "no" ),
                       sd0 = 1
                       )
factors <- factors %>% mutate(
    reps = R / M_CHUNK,
    seed = 16200320 + 1:n()
)
```

This generates a data frame of all our factor combinations.
This is our list of "tasks" (each row of factors).
These tasks have repeats: the "chunks" means we do a portion of each scenario, as specified by our simulation factors, as a process.
This would allow for greater parallelization (e.g., if we had more cores), and also lets us save our work without finishing an entire scenario of, in this case, 1000 iterations.

To set up our simulation we make a little helper method to do one row.
With each row, once we have run it, we save it to disk.
This means if we kill our simulation half-way through, most of the work would be saved.
Our function is then going to either do the simulation (and save the result to disk immediately), or, if it can find the file with the results from a previous run, load those results from disk:

```{r, eval=FALSE}
##### Run the simulation!   #####

scat( "Running simulation in %d chunks\n", nrow(factors) )

tictoc::tic()

# parallel version
safe_run_sim = safely( run_sim )
file_saving_sim = function( chunkNo, seed, ... ) {
    fname = paste0( "results/frags/fragment_", chunkNo, "_", seed, ".rds" )
    res = NA
    if ( !file.exists(fname) ) {
        res = safe_run_sim( chunkNo=chunkNo, seed=seed, ... )
        saveRDS(res, file = fname )
    } else {
        res = readRDS( file=fname )
    }
    return( res )
}
```

Note how we wrap our core `run_sim` method in `safely`; it was crashing very occasionally, and so to make the code more robust, we wrapped it so we could see any error messages.

We next run the simulation.
We shuffle the rows of our task list so that which process gets what task is randomized.
If some tasks are much longer (e.g., due to larger sample size) then this will get balanced out across our processes.

We have an `if-then` structure to easily switch between parallel and nonparallel code.
This makes debugging easier: when running in parallel, stuff printed to the console does not show until the simulation is over.
Plus it would be all mixed up since multiple processes are working simultaneously.

This overall structure allows the researcher to delete one of the "fragment" files from the disk, run the simulation code, and have it just do one tiny piece of the simulation.
This means the researcher can insert a `browser()` command somewhere inside the code, and debug the code, in the natural context of how the simulation is being run.

```{r, eval=FALSE}
# Shuffle the rows so we run in random order to load balance.
factors = sample_n(factors, nrow(factors) )

if ( TRUE ) {
    # Run in parallel
    parallel::detectCores()
    
    library(future)
    library(furrr)
    
    #plan(multiprocess) # choose an appropriate plan from future package
    #plan(multicore)
    plan(multisession, workers = parallel::detectCores() - 2 )
    
    factors$res <- future_pmap(factors, .f = file_saving_sim,
                          .options = furrr_options(seed = NULL),
                          .progress = TRUE )
    
} else {
  # Run not in parallel, used for debugging
  factors$res <- pmap(factors, .f = file_saving_sim )
}

tictoc::toc()
```

Our method cleverly loads files in, or generates them, for each chunk.
The seed setting ensures reproducibility.
Once we are done, we need to clean up our results:

```{r, eval=FALSE}
sim_results <- 
    factors %>% 
    unnest(cols = res)

# Cut apart the results and error messages
sim_results$sr = rep( c("res","err"), nrow(sim_results)/2)
sim_results = pivot_wider( sim_results, names_from = sr, values_from = res )

saveRDS( sim_results, file="results/simulation_results.rds" )

cat( "Script finished\n" )
```



### Dynamically making directories

If you are generating a lot of files, then you should put them somewhere.
But where?
It is nice to dynamically generate a directory for your files on fly.
One way to do this is to write a function that will make any needed directory, if it doesn't exist, and then put your file in that spot.
For example, you might have your own version of `write_csv` as:

```{r}
my_write_csv <- function( data, path, file ) {
  
  if ( !dir.exists( here::here( path ) ) ) {
    dir.create( here::here( path ), recursive=TRUE ) 
  }
  write_csv( data, paste0( path, file ) )
}
```

This will look for a path (starting from your R Project, by taking advantage of the `here` package), and put your datafile in that spot.
If the spot doesn't exist, it will make it for you.

