```{r, include =FALSE}
library( purrr )
```


# Error trapping and other headaches

If you have an advanced estimator, or are relying on some package, it is quite possible that every so often your estimate will trigger an error, or give you a NA result, or something similarly bad.
More innocuous, you might have estimators that can generate warnings; if you run 10,000 trials, that can add up to a lot of warnings which can be overwhelming to sort through.
In some cases, these warnings might be coupled with the estimator returning a very off result; it is unclear, in this case, whether we should include that result in our overall performance measures for that estimator.
After all, it tried to warn us!

In this section, we talk about some ways to make your simulations safe and robust, and also discuss some ways to track warnings and include them in peformance measures.
 

## Safe code {#safe_code}


Sometimes if you write a function that does a lot of complex things with
uncertain objects -- i.e. consider a complex estimator using an unstable package not of your own creation -- you can run into trouble where you have a intermittent error.

This can really be annoying; consider the case where your simulation crashes on 1 out of 500 chance: if you run 1000 simulation trials, your program will likely not make it to the end, thus wasting all of your time.
You can write code that can, instead of stopping when it reaches an error, trap the error and move on with the next simulation trial.

To illustrate, consider the following broken function:

```{r, error=TRUE}
my_complex_function = function( param ) {
    
    vals = rnorm( param, mean = 0.5 )
    if ( sum( vals ) > 5 ) {
        broken_code( 4 )
    } else {
        sqrt( sum( vals ) )
    }
}
```

We run it like so:
```{r, error=TRUE}
my_complex_function( 1 )
my_complex_function( 7 )

resu = rerun( 20, my_complex_function( 7 ) )
```

Oh no!  Our function crashes sometimes.
To trap the errors we use the purrr package to make a "safe" function as so:

```{r}
my_safe_function = safely( my_complex_function )
my_safe_function( 7 )
```

There are other function wrappers in this family, such as "possibly":
```{r}
my_possible_function = possibly( my_complex_function, 
                                 otherwise = NA )
my_possible_function( 7 )

as.numeric( rerun( 10, my_possible_function(7) ) )
```

Finally, we can use "quietly" to make warnings and stuff go away:
```{r}
my_quiet_function = quietly( my_complex_function )

my_quiet_function( 1 )
```

This can be especially valuable to control massive amounts of printout in a simulation.  If you have lots of extraneous printout, it can slow down the execution of your code far more than you might think.


### What to do with warnings

Sometimes our analytic strategy might give some sort of warning (or fail altogether).
For example, from the cluster randomized experiment case study we have:

```{r}
set.seed(101012)  # (I picked this to show a warning.)
dat = gen_dat_model( J = 50, n_bar = 100, sigma2_u = 0 )
mod <- lmer( Yobs ~ 1 + Z + (1|sid), data=dat )
```

We have to make a deliberate decision as to what to do about this:
 
 - Keep these "weird" trials?
 - Drop them?

If you decide to drop them, you should drop the entire simulation iteration including the other estimators, even if they worked fine!
If there is something particularly unusual about the dataset, then dropping for one estimator, and keeping for the others that maybe didn't give a warning, but did struggle to estimate the estimand, would be unfair: in the final performance measures the estimators that did not give a warning could be being held to a higher standard, making the comparisons between estimators biased.

If your estimators generate warnings, you should calculate the rate of errors or warning messages as a performance measure.
Especially if you drop some trials, it is important to see how often things are acting pecularly.

The main tool for doing this is the `quietly()` function:
```{r}
quiet_lmer = quietly( lmer )
qmod <- quiet_lmer( Yobs ~ 1 + Z + (1|sid), data=dat )
qmod
```

You then might have, in your analyzing code:
```{r, eval=FALSE}
analyze_data <- function( dat ) {
    
    M1 <- quiet_lmer( Yobs ~ 1 + Z + (1|sid), data=dat )
    message1 = ifelse( length( M1$message ) > 0, 1, 0 )
    warning1 = ifelse( length( M1$warning ) > 0, 1, 0 )

    # Compile our results
    tibble( ATE_hat = coef(M1)["Z"],
            SE_hat = se.coef(M1)["Z"],
            message = message1,
            warning = warning1 )
}
```

Now you have your primary estimates, and also flags for whether there was a convergence issue.
In the analysis section you can then evaluate what proportion of the time there was a warning or message, and then do subset analyses to those simulation trials where there was no such warning.



## Saving files and results

Always save your simulation results to a file.
Simulations are painful and time consuming to run, and you will invariably want to analyze the results of them in a variety of different ways, once you have looked at your preliminary analysis.
We advocate saving your simulation as soon as it is complete.
But there are some ways to do better than that, such as saving as you go.
This can protect you if your simulation occasionally crashes, or if you want to rerun only parts of your simulation for some reason.

### Saving simulations as you go

If you are not sure you have time to run your entire simulation, or you think your computer might crash half way through, or something similar, you can save each chunk you run as you go, in its own file.  You then stack those files at the end to get your final results.
With clever design, you can even then selectively delete files to rerun only parts of your larger simulation---but be sure to rerun everything from scratch before you run off and publish your results, to avoid embarrassing errors.

Here, for example, is a script from a research project examining how one might use post-stratification to improve the precision of an IV estimate.
This is the script that runs the simulation.
Note the sourcing of other scripts that have all the relevant functions; these are not important here.
Due to modular programming, we can see what this script does, even without those detail.

```{r IV_simulation_driver, eval=FALSE}
source( "pack_simulation_functions.R" )

if ( !file.exists("results/frags" ) ) {
    dir.create("results/frags")
}

# Number of simulation replicates per scenario
R = 1000

# Do simulation breaking up R into this many chunks
M_CHUNK = 10

###### Set up the multifactor simulation #######

# chunkNo is a hack to make a bunch of smaller chunks for doing parallel more
# efficiently.
factors = expand_grid( chunkNo = 1:M_CHUNK,
                       N = c( 500, 1000, 2000 ),
                       pi_c = c( 0.05, 0.075, 0.10 ),
                       nt_shift = c( -1, 0, 1 ),
                       pred_comp = c( "yes", "no" ),
                       pred_Y = c( "yes", "no" ),
                       het_tx = c( "yes", "no" ),
                       sd0 = 1
                       )
factors <- factors %>% mutate(
    reps = R / M_CHUNK,
    seed = 16200320 + 1:n()
)
```

This generates a data frame of all our factor combinations.
This is our list of "tasks" (each row of factors).
These tasks have repeats: the "chunks" means we do a portion of each scenario, as specified by our simulation factors, as a process.
This would allow for greater parallelization (e.g., if we had more cores), and also lets us save our work without finishing an entire scenario of, in this case, 1000 iterations.

To set up our simulation we make a little helper method to do one row.
With each row, once we have run it, we save it to disk.
This means if we kill our simulation half-way through, most of the work would be saved.
Our function is then going to either do the simulation (and save the result to disk immediately), or, if it can find the file with the results from a previous run, load those results from disk:

```{r, eval=FALSE}
safe_run_sim = safely( run_sim )
file_saving_sim = function( chunkNo, seed, ... ) {
    fname = paste0( "results/frags/fragment_", chunkNo, "_", seed, ".rds" )
    res = NA
    if ( !file.exists(fname) ) {
        res = safe_run_sim( chunkNo=chunkNo, seed=seed, ... )
        saveRDS(res, file = fname )
    } else {
        res = readRDS( file=fname )
    }
    return( res )
}
```

Note how we wrap our core `run_sim` method in `safely`; it was crashing very occasionally, and so to make the code more robust, we wrapped it so we could see any error messages.

We next run the simulation.
We shuffle the rows of our task list so that which process gets what task is randomized.
If some tasks are much longer (e.g., due to larger sample size) then this will get balanced out across our processes.

We have an `if-then` structure to easily switch between parallel and nonparallel code.
This makes debugging easier: when running in parallel, stuff printed to the console does not show until the simulation is over.
Plus it would be all mixed up since multiple processes are working simultaneously.

This overall structure allows the researcher to delete one of the "fragment" files from the disk, run the simulation code, and have it just do one tiny piece of the simulation.
This means the researcher can insert a `browser()` command somewhere inside the code, and debug the code, in the natural context of how the simulation is being run.

```{r, eval=FALSE}
# Shuffle the rows so we run in random order to load balance.
factors = sample_n(factors, nrow(factors) )

if ( TRUE ) {
    # Run in parallel
    parallel::detectCores()
    
    library(future)
    library(furrr)
    
    #plan(multiprocess) # choose an appropriate plan from future package
    #plan(multicore)
    plan(multisession, workers = parallel::detectCores() - 2 )
    
    factors$res <- future_pmap(factors, .f = file_saving_sim,
                          .options = furrr_options(seed = NULL),
                          .progress = TRUE )
    
} else {
  # Run not in parallel, used for debugging
  factors$res <- pmap(factors, .f = file_saving_sim )
}

tictoc::toc()
```

Our method cleverly loads files in, or generates them, for each chunk.
The seed setting ensures reproducibility.
Once we are done, we need to clean up our results:

```{r, eval=FALSE}
sim_results <- 
    factors %>% 
    unnest(cols = res)

# Cut apart the results and error messages
sim_results$sr = rep( c("res","err"), nrow(sim_results)/2)
sim_results = pivot_wider( sim_results, names_from = sr, values_from = res )

saveRDS( sim_results, file="results/simulation_results.rds" )
```



### Dynamically making directories

If you are generating a lot of files, then you should put them somewhere.
But where?
It is nice to dynamically generate a directory for your files on fly.
One way to do this is to write a function that will make any needed directory, if it doesn't exist, and then put your file in that spot.
For example, you might have your own version of `write_csv` as:

```{r}
my_write_csv <- function( data, path, file ) {
  
  if ( !dir.exists( here::here( path ) ) ) {
    dir.create( here::here( path ), recursive=TRUE ) 
  }
  write_csv( data, paste0( path, file ) )
}
```

This will look for a path (starting from your R Project, by taking advantage of the `here` package), and put your data file in that spot.
If the spot doesn't exist, it will make it for you.




### Loading and combining files of simulation results

Once your simulation files are all generated, the following code will stack them all into a giant set of results, assuming all the files are themselves data frames stored in RDS objects.
This function will try and stack all files found in a given directory; for it to work, you should ensure there are no other files stored there.


```{r, eval=FALSE}
load.all.sims = function( filehead="results/" ) {
  
  files = list.files( filehead, full.names=TRUE)
  
  res = map_df( files, function( fname ) {
    cat( "Reading results from ", fname, "\n" )
    rs = readRDS( file = fname )
    rs$filename = fname
    rs
  })
  res
}
```

You would use as so:
```{r, eval=FALSE}
results = load.all.sims( filehead="raw_results/" )
```



