# Design, analysis, and presentation of simulation results

## Designing the simulation experiment 

- The primary limitation of simulation studies is __generalizability__.
- Choose conditions that allow you to relate findings to previous work.
- Err towards being comprehensive.
    - The goal should be to build an understanding of the major moving parts.
    - Presentation of results can always be tailored to illustrate trends.
- Explore breakdown points (e.g., what sample size is too small for applying a given method?).

## Choosing parameter levels 

Four possible strategies: 

1. Vary a parameter over its entire range (or nearly so).
2. Choose parameter levels to represent realistic practical range.
    - Empirical justification based on systematic reviews of applications
    - Or at least informal impressions of what's realistic in practice
3. Vary nuisance parameters (at least a little) to test sensitivity of results.
4. Choose parameters to emulate one important application.

## Presentation

Your results have finished running...what now?

- Understand the effects of all of the factors manipulated in the simulation
- Develop evidence that addresses your research questions
    
Three approaches to analysis and presentation:

1. Tabulation
2. Visualization
3. Modeling

## Tabulation 

- Traditionally, simulation study results are presented in big tables. Tables are fine if...
    - they involve only a few numbers, and a few targeted comparisons 
    - it is important to report _exact_ values for some quantities

- But simulations usually produce lots of numbers, and involve making lots of comparisons.
    - relative performance of alternative estimators
    - performance under different conditions for the data-generating model

- Exact values for bias/RMSE/type-I error are not usually of interest.

- It is often more useful and insightful to present results in graphs (Gelman, Pasarica, & Dodhia, 2002).

## Visualization

Visualization should nearly always be the first step in analyzing simulation results.

This often requires creating a _BUNCH_ of graphs to look at different aspects of the data.

Helpful tools/concepts:

- Boxplots are often useful for depicting range and central tendency across many combinations of parameter values. 
- Color, shape, line type to encode different factors
- Small multiples (faceting) to encode further factors (e.g., varying sample size)

### Example 1 

- Bias of biserial correlation estimate from an extreme groups design
- $96 \times 2 \times 5 \times 5$ factorial design (true correlation, cut-off type, cut-off percentile, sample size)

```{r, echo=FALSE, warning=FALSE, fig.width = 10, fig.height = 4}
load("data/d2r results.rData")
allResults$n <- ordered(allResults$n)
allResults$p.inv <- allResults$p1
allResults$p1 <- ordered(allResults$p1, 
                labels = paste("p1 = 1/",unique(allResults$p1), sep=""))
allResults$fixed <- ordered(allResults$fixed, levels=c("TRUE","FALSE"), 
                labels = c("Fixed percentiles","Sample percentiles"))

r_F <- droplevels(subset(allResults, stat=="r.i" & design=="Extreme Group"))
levels(r_F$fixed) <- c("Pop. cutoff","Sample cutoff")
r_F$bias <- r_F$mean - r_F$rho
r_F$bias.sm <- r_F$mean.sm - r_F$rho
r_F$rmse <- sqrt(r_F$bias^2 + r_F$var)

library(ggplot2)
ggplot(r_F, aes(rho, bias, linetype = n)) +
  geom_smooth(method="loess", se=FALSE, color = "black") + 
  facet_grid(fixed ~ p1) + theme_bw() +
  labs(linetype = "n") +
  scale_y_continuous(name=expression(Bias(r[eg]))) + 
  scale_x_continuous(name=expression(rho))
```

Source: Pustejovsky, J. E. (2014). Converting from d to r to z when the design uses extreme groups, dichotomization, or experimental control. Psychological Methods, 19(1), 92-112.

### Example 2 

- Type-I error rates of small-sample corrected F-tests based on cluster-robust variance estimation in meta-regression
- Comparison of 5 different small-sample corrections
- Complex experimental design, varying
    - sample size ($m$)
    - dimension of hypothesis ($q$)
    - covariates tested
    - degree of model mis-specification

```{r, echo=FALSE, fig.height=5, fig.width=9}
load("data/RVE_simulation.Rdata")
results <- results_large_m
results <- within(results, {
  type <- substr(contrast,1,1)
  q <- as.numeric(substr(contrast,3,3))
  p <- ifelse(type=="O",q + 1,6)  
  q_lab <- factor(q)
  levels(q_lab) <- paste("q =", levels(q_lab)) 
  testname <- factor(test, levels = c("Chi-sq (Uncorrected)","Chi-sq","Naive F",
                                      "Fay-Cornelius 2","Cai-Hayes 1","T-sq Z","T-sq B","T-sq A",
                                      "Fay-Cornelius 1","Cai-Hayes 2","Cai-Hayes 3",
                                      "Satterthwaite 1","Satterthwaite 2","Satterthwaite 3","Satterthwaite 4",
                                      "PW-eigen","Zhang-eigen"))
  levels(testname)[which(levels(testname)=="Chi-sq")] <- "Chi-sq (BRL)"
  levels(testname)[which(levels(testname) %in% c("Fay-Cornelius 2","Cai-Hayes 1"))] <- c("EDF","EDT")
  levels(testname)[which(levels(testname) %in% c("T-sq Z","T-sq B","T-sq A"))] <- c("T^2 Z","T^2 B","T^2 A")
})

iterations <- 5000
MC_CI <- qnorm(0.975) * sqrt(0.05 * 0.95 / iterations)

test_select <- c("EDT","EDF","T^2 A", "T^2 B","T^2 Z")
m_select <- c(10, 20, 40, 80)

ggplot(subset(results, testname %in% test_select & m %in% m_select & q < 5),
       aes(testname, p05, fill = testname)) + 
  geom_boxplot() + 
  facet_grid(q ~ m, scales = "free_y",labeller = "label_both") + 
  scale_x_discrete(labels = abbreviate) + 
  labs(x = NULL, y = "Type I error", fill = "Test") + 
  geom_hline(yintercept= 0.05) + 
  geom_hline(yintercept= 0.05 + MC_CI, linetype = "dashed") +
  theme_bw()

```

Source: Tipton, E., & Pustejovsky, J. E. (2015). Small-sample adjustments for tests of moderators and model fit using robust variance estimation in meta-regression. _Journal of Educational and Behavioral Statistics, 40_(6), 604-634.

### Example 3 (Pustejovsky & Swan, 2014)

- Coverage of parametric bootstrap confidence intervals for momentary time sampling data
- Compare maximum likelihood estimators to posterior mode (penalized likelihood) estimators of prevalence
- 2-dimensional parameter space:
    - prevalence (19 levels)
    - incidence (10 levels)
- Sample size (15 levels)

Source: Pustejovsky, J. E., & Swan, D. M. (2015). Four methods for analyzing partial interval recording data, with application to single-case research. _Multivariate Behavioral Research, 50_(3), 365-380.

```{r, echo=FALSE, fig.height=5.5, fig.width = 10}

library(plyr)

load("data/MTS bootstrap performance.Rdata")

MTS_results <- BSresults

breaks_coverage <- c(0, 0.925, 0.94, 0.96, 0.975, 1)
labels_coverage <- c("0-92.5%", "92.5-94%", "94-96%", "96-97.5%", "97.5-100%")
coverage_colors <- c("0-92.5%" = "pink", "92.5-94%" = "yellow" , "94-96%" = "green", "96-97.5%" = "blue", "97.5-100%" = "purple")

coverage_smoother <- function(results){
  pcoverage_model <- loess(pcoverage ~ phi + zeta, data = results, span = 0.25)
  zcoverage_model <- loess(zcoverage ~ phi + zeta, data = results, span = 0.25)
  pcoverage_smooth <- predict(pcoverage_model, newdata = results)
  zcoverage_smooth <- predict(zcoverage_model, newdata = results)
  
  return(cbind(results, pcoverage_smooth, zcoverage_smooth))
}

MTS_coverage <- ddply(MTS_results, .fun = coverage_smoother, .(K_intervals, k_priors, theta))
MTS_coverage2 <- MTS_coverage
MTS_coverage2$phi <- 1 - MTS_coverage2$phi
MTS_coverage <- rbind(MTS_coverage, MTS_coverage2)
MTS_coverage$pcoverage_smooth <- ifelse(MTS_coverage$pcoverage_smooth > 1, 1, MTS_coverage$pcoverage_smooth)

MTS_coverage$pcoverage_cut <- cut(MTS_coverage$pcoverage, breaks = breaks_coverage,
                              labels = labels_coverage, include.lowest = TRUE)
MTS_coverage$pcoverage_cut_smooth <- cut(MTS_coverage$pcoverage_smooth, breaks = breaks_coverage,
                                     labels = labels_coverage, include.lowest = TRUE)

MTS_coverage$zcoverage_smooth <- ifelse(MTS_coverage$zcoverage_smooth > 1, 1, MTS_coverage$zcoverage_smooth)

MTS_coverage$zcoverage_cut <- cut(MTS_coverage$zcoverage, breaks = breaks_coverage,
                              labels = labels_coverage, include.lowest = TRUE)
MTS_coverage$zcoverage_cut_smooth <- cut(MTS_coverage$zcoverage_smooth, breaks = breaks_coverage,
                                     labels = labels_coverage, include.lowest = TRUE)


qplot(phi, zeta, fill = pcoverage_cut_smooth, 
      geom = "tile",
      data = subset(MTS_coverage, theta == Inf & K_intervals >= 40)) +
  facet_wrap(~K_intervals, ncol = 4, scales = "free_y") +
  scale_y_continuous(breaks=seq(.1, .50, .1)) + 
  scale_x_continuous(breaks=seq(.1, 1, .1)) +
  scale_fill_manual(values = coverage_colors) + 
  labs(x = "Prevalence", y = "Incidence", fill = "Coverage") + theme_bw()+ 
  theme(axis.text.x = element_text(angle=45, hjust = 1), legend.position = "bottom")
```

```{r, echo=FALSE, fig.height=5.5, fig.width = 10}
qplot(phi, zeta, fill = pcoverage_cut_smooth, 
      geom = "tile",
      data = subset(MTS_coverage, theta == 10 & K_intervals >= 40)) +
  facet_wrap(~K_intervals, ncol = 4, scales = "free_y") +
  scale_y_continuous(breaks=seq(.1, .50, .1)) + 
  scale_x_continuous(breaks=seq(.1, 1, .1)) +
  scale_fill_manual(values = coverage_colors) + 
  labs(x = "Prevalence", y = "Incidence", fill = "Coverage") + theme_bw()+
  theme(axis.text.x = element_text(angle=45, hjust = 1), legend.position = "bottom")
```

## Modeling

Simulations are designed experiments, often with a full factorial structure

Helpful tools:

>- ANOVA can be useful for understanding major sources of variation in simulation results (e.g., identifying which factors have negligible/minor influence on the bias of an estimator).
>- Smoothing (e.g., local linear regression) over continuous factors 


### Example 1

- Bias of biserial correlation estimate from an extreme groups design
- $96 \times 2 \times 5 \times 5$ factorial design (true correlation, cut-off type, cut-off percentile, sample size)

```{r, warning=FALSE}
anova_table <- aov(bias ~ rho * p1 * fixed * n, data = r_F)
summary(anova_table)
```

```{r, warning=FALSE}
library(lsr)
# etaSquared(anova_table)
```

## Presentation

- Present selected results that clearly illustrate the main findings from the study and anything unusual/anomolous.
- In the text of your write-up, include examples that make specific numerical comparisons. 
- Include supplementary materials containing
    - additional figures and analysis
    - complete simulation results 
    - reproducible code for running the simulation and doing the analysis
