---
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup-exp-design, include=FALSE}
library( tidyverse )
library( purrr )
options(list(dplyr.summarise.inform = FALSE))
theme_set( theme_classic() )

### Code for the running examples
source( "case_study_code/r_bivariate_Poisson.R" )
source( "case_study_code/r_and_z.R" )
source( "case_study_code/evaluate_CIs.R" )

source( "case_study_code/clustered_data_simulation.R" )

dat <- gen_cluster_RCT( n=5, J=3, p=0.5, 
                      gamma_0=0, gamma_1=0.2, gamma_2=0.2,
                      sigma2_u = 0.4, sigma2_e = 1,
                      alpha = 0.5 )

```

# Designing multifactor simulations {#designing-multifactor-simulations}

Thus far, we have created code that will run a simulation for a single combination of parameter values.
In practice, simulation studies typically examine a range of different values, including varying the levels of the focal parameter values, auxiliary parameters, sample size, and possibly other design parameters, to explore a range of different scenarios.
We either want reassurance that our findings are general, or we want to understand what aspects of the context affect the performance of the estimator or estimators we are studying.
A single simulation gives us no hint as to either of these questions.
It is only by looking across a range of settings that we can hope to understand trade-offs, general rules, and limits.
Let's now look at the remaining piece of the simulation puzzle: the study's experimental design. 

Simulation studies often take the form of __full factorial__ designed experiments.
In full factorials, each factor (a particular knob a researcher might turn to change the simulation conditions) is varied across multiple levels, and the design includes _every_ possible combination of the levels of every factor. 
The multi-factor aspect of a simulation is incredible important.
It can take us from an overly narrow exploration to one that has broader significance.
As @little2013praise puts it:

> Good simulation studies are not given the respect they deserve. Often the design is perfunctory and simplistic, neglecting to attempt a factorial experimental design to cover the relevant sample space, and results are over-generalized. Well designed simulation studies with realistic sample sizes are an antidote to a ﬁxation on asymptotics and a useful tool for assessing calibration.

In Section \@ref(simulating-across-multiple-factors), we demonstrated how to represent a full factorial design as a list of the factors and levels to be explored, and then how to create a dataset containing parameter values for every combinations of levels of all the factors. 
In that example, we designed a simulation study to examine the performance of confidence intervals for Pearson's correlation coefficient under a bivariate Poisson distribution. 
We arrived at a $3 \times 3 \times 2 \times 8$ factorial design, varying the sample size across three levels, the mean of one variable across three levels, the ratio of means across two levels, and the true correlation between variables across eight levels.

But how do we go about choosing parameter values to examine?
Choosing which parameters to use is a central part of good simulation design because the primary limitation of simulation studies is always their _generalizability_.
On the one hand, it is difficult to extrapolate findings from a simulation study\index{simulation study} beyond the set of simulation conditions that were examined. On the other hand, it is often difficult or impossible to examine the full space of all possible parameter values, except for very simple problems.
Even in the relatively straightforward Pearson correlation simulation, we have four factors and the last three could each take on an infinite number of possible levels.
How do we come up with a defensible set of levels to examine?

## Choosing factors and parameter combinations

As far as we know, there is no generic algorithm for choosing which parameters to vary in a simulation design, nor for choosing what levels to use for each parameter.
The particular answers to these questions will always depend on the problem you are studying, the focus of your research questions, and the availability of any relevant theoretical results. 
Because choices of simulation conditions need to be made in the context of the problem or model that you are studying, it is difficult to identify valid but acontextual rules.
Instead, we can only offer broad principles and practical advice, informed by our experience conducting and reading simulation studies.

Broadly, we generally try to vary parameters that we believe matter, or that we think other people will believe matter.
The first is so we can learn for ourselves.
The second is to build our case.
We see three distinct considerations that are useful to weigh when choosing factors and levels:

1. __Relation to existing research__. Especially for research simulations, it is important to be able to relate your findings to previous research. This suggests that you should select parameter levels that make this possible, such as by looking at sample sizes similar to those examined in previous studies. That said, previous simulation studies are not always perfect (in truth, there are a lot of quite bad ones out there!), and so prior work should not generally be your sole guide or justification.

2. __Comprehensiveness__. Generally, t is usually better to err on the side of being more comprehensive. You learn more by looking at a broader range of conditions, and you can always boil down your results to a more limited set of conditions for purposes of presentation.

3. __Pushing boundaries__. One of the most useful potential contributions of research simulations is to explore the breakdown points of a method (e.g., what sample size is too small for an estimator to be trusted?) rather than focusing only on conditions where a method might be expected to work well. Pushing the boundaries and identifying conditions where estimation methods break will help you to provide better guidance for how the methods should be used in practice.

Once you have identified your parameters, you then have to decide on the levels of the parameter you will include in the simulation.
There are at least three strategies you might take:

1. Vary a parameter over as much of its range as you can.
2. Choose parameter levels to represent a realistic practical range. These ranges would ideally be empirically justified based on systematic reviews of prior applications. Lacking such empirical evidence, you may need to rely on informal impressions of what is realistic in practice.
3. Choose parameters to emulate an important application or context.

Of these choices, option (1) is the most general but also the most computationally intensive.
In the ideal, option (2) focuses attention on what is most relevant to a practitioner.
Option (3) is usually coupled with a subsequent applied data analysis, and in this case the simulation is often used to enrich that analysis.
In particular, if the simulation shows the methods work for data with the given form of the target application, people may be more willing to believe the application's findings.

Regardless of how you select your primary parameters, you should also vary nuisance parameters (at least a little) to test the sensitivity of your results to these other aspects.
Of course, no simulation will ever be fully generalizable.
Still, it is worth considering how to design a simulation to avoid obvious limitations that a critic might identify as a basis for dismissing your findings.

Using a more comprehensive design will generally increase the amount of computing required for a simulation. However, this can be tempered by reducing the number of replications per scenario.
For example, suppose you were initially planning on doing 1000 simulations per scenario, but then you realize  that you need to incorporate an additional factor.
After thinking about it, you suspect that this factor will not actually matter but you worry that other researchers might find it important.
To address this potential concern, you could include the factor in your simulation design, say with four levels, but then reduce the number of replications to 250 simulations per scenario.
The total work remains the same.
When analyzing the final simulation you would first verify you do not see trends along this new factor, and then marginalize out that factor in your summaries of results.
Marginalizing out a factor (i.e., averaging your performance metrics across the additional factor) is a powerful technique of making a claim about how your methods work _on average_ across a _range_ of scenarios, rather than for a specific scenario.
We will discuss it further in Chapter \@ref(presentation-of-results).

In summary, when we are thinking about how to set up a simulation design, we always try to keep the following principles in mind:

- The primary limitation of simulation studies is generalizability.
- Choose conditions that allow you to relate your findings to features of empirical data and to previous methodological work.
- Err towards being comprehensive, with the goal of building an understanding of the major moving parts.
- Explore breakdown points and boundary conditions (e.g., what sample size is too small for applying a given method?).

Finally, we almost always expect to add and subtract from the set of simulation factors after examining an initial set of results. 
In our experience, it is quite rare to nail the choice of parameters on the first pass.
In revising an initial simulation design, the goal should never be to make a particular estimator look better or to conceal problems with the performance of a method. 
Rather, it should be to identify the set of factors and parameter settings that are most relevant and most useful for investigating the questions at issue. 
After all, the final design that we arrive at needs to be well justified and defensible.


## Case Study: A multifactor evaluation of cluster RCT estimators

To bring the multifactor simulation to life, let us return to the case study of comparing three ways to analyze a cluster randomized trial that we presented in Section \@ref(case-cluster).
In our original setup, we wrote code to generate cluster randomized trial data where a cluster's size could be correlated its average treatment effect.
In Chapter \@ref(clusterRCTperformance), we looked across a variety of performance criteria so we could see how the estimators compared for any given scenario we wanted.

So far, we have only examined a single scenario at a time.
But how do our findings generalize? Under what conditions do the various estimation methods perform better or worse? 
To answer these questions, we need to extend to a multifactor simulation to _systematically_ explore how our three estimators behave across a range of contexts.
Happily, the modular functions that we have designed make it relatively straightforward to explore a range of scenarios by calling our simulation function over and over, using the tools of this chapter.
We now illustrate how to extend this simulation to a full multifactor design.
Then in Chapter \@ref(building-good-visualization), we will continue this running example by examining how to analyze the results of the multifactor design.


### Choosing parameters for the Clustered RCT
\index{example!cluster randomized trial}

We begin by identifying some potential research questions suggested by our preliminary exploration.
Regarding bias, we noticed in our initial simulation that Linear Regression targets a person-weighted average effect, so it would be considered biased for the cluster-average treatment effect.
But how large is this bias in practice, and how much does bias change as we change the cluster-size by impact relationship?
Considering precision, we saw that Linear Regression has a higher standard error than the other estimators.
Is this a general pattern?  If not, are there contexts where linear regression will have a lower standard error than the others?
Further, we originally thought that aggregation would lose information because smaller clusters would have the same weight as larger clusters, but be more imprecisely estimated.
Were we wrong? Or perhaps if cluster size was even more variable, aggregation might do worse and worse.
Finally, the estimated SEs for all three methods all appeared to be good, although they were rather variable, relative to the true SE.
We might then ask, are the standard errors always the right size, on average?  Will the estimated SEs fall apart (i.e., be far too large or far too small) in some contexts?  If so, which ones?

To answer all of these questions we need to more systematically explore the space of data-generating conditions.
Our data-generating function has many knobs to turn.
In particular, our data-generating process\index{data-generating process} will produce artificial cluster-randomized experiments where we can vary any of the following features:

 - the number of clusters;
 - the proportion of clusters that receive treatment;
 - the average cluster size and degree of variation in cluster sizes;
 - how much the average impact varies across clusters, and how strongly that is connected to cluster size;
 - how much the cluster intercepts vary (degree of cross-cluster variation); and
 - the degree of residual variation.
 
Manipulating all of these factors would lead to a huge and unwieldy number of simulation conditions to evaluate.
Before proceeding, we reflect on our research questions, speculate as to what is likely to matter, and then consider varying the following:

 - Number of clusters: Do cluster-robust SEs work with fewer clusters?
 - Average cluster size: Does the number of students/cluster matter?
 - Variation in cluster size: Do varying cluster sizes cause bias or break things?
 - Correlation of cluster size and cluster impact: Will correlation cause bias?
 - Cross cluster variation: Does the amount of cluster variation matter?
 
When selecting factors to manipulate, it is important to ensure the each factor is isolated, so that changing one of them should not change other aspects of the data-generating process that might impact performance.
For example, if we simply added more cross-cluster variation by directly increasing the random effects for the clusters, the total variation in the outcome will also increase.
If we then see that the performance of an estimator deteriorates as variation increases, we have a confound: is the cross-cluster variation causing the problem, or is it the total variation?
To avoid this confound, we should vary cluster variation while holding the total variation fixed; this is why we use the ICC parameterization, as discussed in Section \@ref(case-cluster).

Given our research questions and the way we parameterize the DGP\index{data-generating process}, we end up with the following factors and levels:

```{r CRT_factors}
crt_design_factors <- list(
  n_bar = c( 20, 80, 320 ),
  J = c( 5, 20, 80 ),
  ATE = c( 0.2 ),
  size_coef = c( 0, 0.2 ),
  ICC = c( 0, 0.2, 0.4, 0.6, 0.8 ),
  alpha = c( 0, 0.5, 0.8 )
)
```

The ATE factor only has one level. We could later expand this if we wanted to, but based on theoretical concerns we are fairly confident the ATE will not impact our research questions, so we leave it as it is, set at a value we would consider plausible in real life.

### Redundant factor combinations

There is some redundancy in the parameter combinations that we have selected.
In particular, if `size_coef` is nonzero, but `alpha` is 0, then the cluster size will not impact the cluster average ATE because there is no variation in cluster size.
We could drop one of the redundant conditions to save some simulation runs---in essence we are running the same simulation for `alpha=0` two times, once for each `size_coef` value, for each level of `ICC`, `n_bar` and `J`.
However, doing so would mean that our simulation is no longer fully crossed.
We will leave the redundant conditions in as a sanity check for our results. 
We know we should get the same results; if we do not, then we either have uncontrolled uncertainty in our simulation or an error in the code.
Because computation is cheap, then keeping the fully crossed structure will make for easier analysis.
Otherwise our experiment would become imbalanced, and so when we average performance across some factors to see the effect of others, we might end up with surprising or even misleading results.

### Running the simulations

We will run our cluster RCT simulation using the same code pattern as we used with the Pearson correlation simulations.
Because we are not exactly sure which performance metrics we will want to use, we will save the individual replications and then calculate performance metrics after generating the simulation results.
That is, we will use the outside strategy.

We first make a table of scenarios:
```{r}
params <- 
  expand_grid( !!!crt_design_factors ) %>%
  mutate(
    seed = 20200320 + 17 * 1:n()
  )
```

To allow reproducibility, we specify a different seed for each scenario just to avoid anything confusing about shared randomness across scenarios (see Section \@ref(seeds-and-pseudo-RNGs) for further discussion).
We then run the simulation 1000 times each for scenario, then unnest to get our final data:

```{r run_the_CRT_simulation, eval=FALSE}
params$res <- pmap(params, .f = run_CRT_sim, reps = 1000 )
res <- unnest(params, cols=data)
saveRDS( res, file = "results/simulation_CRT.rds" )
```


```{r secret_run_full_CRT, include=FALSE}
if ( !file.exists( "results/simulation_CRT.rds" ) ) {
  source( here::here( "case_study_code/clustered_data_simulation_runner.R" ) )
} else {
  res = readRDS("results/simulation_CRT.rds")
}
```

Normally, we would speed up these calculations using parallel processing\index{parallel processing}; we discuss how to do so in Chapter \@ref(parallel-processing).
Even under parallel processing, this simulation took overnight to run.
Our final results look like this:

```{r, echo = FALSE}
res
```

We have a lot of rows of data!


### Calculating performance metrics

Our next step is to group the results by the simulation factors and calculate the performance metrics across the replications of each simulation condition.
Here we calculate our primary performance measures by hand:

```{r, messages=FALSE}
res <- readRDS( file = "results/simulation_CRT.rds" )

sres <- 
  res %>% 
  group_by( n_bar, J, ATE, size_coef, ICC, alpha, method ) %>%
  summarise( 
    bias = mean(ATE_hat - ATE),
    SE = sd( ATE_hat ),
    RMSE = sqrt( mean( (ATE_hat - ATE )^2 ) ),
    ESE_hat = sqrt( mean( SE_hat^2 ) ),
    SD_SE_hat = sqrt( sd( SE_hat^2 ) ),
    power = mean( p_value <= 0.05 ),
    R = n(),
    .groups = "drop"
  )
```

If we want MCSEs (as we usually do), then we could do that by hand or use the `simhelpers` package as so:
```{r}
library( simhelpers )

sres <- 
  res %>% 
  group_by( n_bar, J, ATE, size_coef, ICC, alpha, method ) %>%
  summarise( 
    calc_absolute( 
      estimates = ATE_hat, true_param = ATE, 
      criteria = c("bias","stddev","rmse")
    ),
    calc_relative_var( 
      estimates = ATE_hat, var_estimates = SE_hat^2,
      criteria = "relative bias"
    ) 
  ) %>%
  rename( SE = stddev, SE_mcse = stddev_mcse ) %>%
  dplyr::select( -K_absolute, -K_relvar ) %>%
  ungroup()

glimpse( sres )
```


## Summary

In this chapter, we have moved from isolated simulation scenarios to fully specified multifactor simulation designs.
By systematically varying multiple features of a data-generating process while holding others fixed, we can explore how the performance of an estimator depends on the context in which it is used.
By thinking of simulations as designed experiments, we gain a principled framework for choosing parameters and curating a set of scenarios to explore.
The result is a rich collection of simulation output that can capture bias, variability, uncertainty estimation, or testing behavior across a wide range of plausible conditions.

However, the volume and complexity of systematic simulation results also create new challenges for interpretation.
With many factors, levels, and performance measures, it is no longer possible to understand results simply by inspecting a table of results.
In the next few chapters, we therefore turn to the problem of analyzing and presenting simulation results, with an emphasis on graphical approaches that clarify how performance varies across conditions.
Through further analysis, we will seek to identify systematic trends in estimator performance, understand the generality of identified patterns, and identify trade-offs between different methods.


## Exercises

### Brown and Forsythe redux

Take another look at Table [5.1](case-ANOVA.html#tab:BF-Scenarios), which is excerpted from @brown1974SmallSampleBehavior. 
Create a tibble with one row for each of the 20 scenarios that they evaluated. 
Then create a function for running the full simulation process (see Exercise \@ref(Welch-simulation)).
Use `pmap()` or `evaluate_by_row()` to run simulations of all 20 scenarios and reproduce the results in Table [5.2](case-ANOVA.html#tab:BF-table1) of Chapter \@ref(case-ANOVA).

### Meta-regression

Exercise \@ref(meta-regression-DGP) described the random effects meta-regression model. 
List the focal, auxiliary, and structural parameters of this model, and propose a set of design factors to use in a multifactor simulation of the model.
Create a list with one entry per factor, then create a dataset with one row for each simulation context that you propose to evaluate.


### Comparing the trimmed mean, median and mean {#exercise:trimmed-mean}

In this exercise, you will write a simulation to compare several different
estimators of a common parameter.
In particular, you will compare the  mean, trimmed mean, and median as estimators of the center of a symmetric distribution (such that the mean and median parameters are identical).  
To do this, you should break building this simulation evaluation down into functions for each component of the simulation. 
This will allow you to extend the same framework to more complicated simulation studies.
This extended exercise illustrates how methodologists might compare different estimation strategies, as you might see in the "simulation" section of a stats paper.

As the data-generation function, use a scaled $t$-distribution so that the standard deviation will always be 1 but will have different fatness of tails (high chance of outliers):

```{r}
sres %>%
  group_by( method ) %>%
  summarise( mean_bias = mean( bias ),
             mean_SE = mean( SE ),
             mean_rmse = mean( rmse ) )
```

Even with this very crude summary of the results, we see tantalizing hints of differences between the methods.
Linear Regression does seem biased, on average, across the scenarios considered.
But so does multilevel modeling---we did not see bias before, but apparently for some scenarios it is biased as well.
Overall SE and RMSE\index{performance measures!RMSE} seems roughly the same across scenarios, and it is clear that uncertainty trumps bias, at least on average.
However, this pattern could be driven by the scenarios with few clusters that are small in size.
In order to understand the full story of what is going on, we need to dig further into the results.




