```{r, include=FALSE}
library( tidyverse )
conflicted::conflicts_prefer(dplyr::select)
conflicted::conflicts_prefer(dplyr::filter)

rm( list = ls() )

theme_set( theme_minimal() )
```


# Case study: Comparing different estimators

<!-- LWM: I am not sure we need this chapter anymore.  I like it as a nice stand-alone bit talking about building up a viz, but it seems redundant with the rest of our text? To discuss. -->

Pretend, for a moment, that we have "invented" the trimmed mean and want to demonstrate its utility.
We might do this by showing how it performs in a variety of scenarios, comparing it to the mean and median.
(In fact, you may have done this in Exercise @exr-trimmed-mean.)
This activity is  analogous to what is often done in a simulation methodology paper: invent something, and then demonstrate its utility via simulation.

In this chapter, we walk through how we might analyze the results from such a simulation.
In our simulation, we have three factors: sample size, a degrees of freedom to control the thickness of the tails (thicker tails means higher chance of outliers), and a degree of skew (how much the tails tend to the right vs. the left).
For our data-generation function we use a scaled and skewed $t$-distribution parameterized so the standard deviation will always be 1 and the mean will always be 0, with parameters to control the skew and tail thickness.
See the [attached code file](case_study_code/trimmed_mean_simulation.R) for the simulation. 

Here are the results from our final multi-factor simulation (we ran 1000 trials per scenario):

```{r}
results <- read_rds( here::here( "results/skewed_t_simulation.rds" ) ) %>%
  dplyr::select( -seed )
results
```

## Investigating precision

Our first question is how sample size impacts our different estimators' precision.
We plot:

```{r}
#| fig-height: 4
#| fig-width: 10

ns <- unique( results$n )

ggplot(results) + 
  aes(x=n, y=SE, col=estimator) +
  facet_grid( skew ~ df0 , labeller = "label_both" ) +
  geom_line() + geom_point() +
  scale_x_log10( breaks=ns ) + 
  scale_color_viridis_d()
```

Our plot doesn't show differences clearly because all the SEs goes to zero as $n$ increases.
One move when we see strong trends like this is to log our outcome or otherwise re-scale the variables. 
Using log-10 scales for both axes makes it clear that relative differences in precision are nearly constant across sample sizes.

```{r}
#| fig-height: 4
#| fig-width: 10

ggplot( results ) +
  aes( x=n, y=SE, col=estimator ) +
  facet_grid( skew ~ df0 , labeller = "label_both" ) +
  geom_line() + geom_point() +
  scale_x_log10( breaks=ns ) +
  scale_y_log10() + 
  scale_color_viridis_d()
```

One step better would be to re-scale based on our knowledge of standard errors.
If we scale by the square root of sample size, we should get horizontal lines.
Here are our results:

```{r}
results <- mutate( results, scaleSE = SE * sqrt(n) )
```

```{r}
#| fig-height: 4
#| fig-width: 10

ggplot( results ) +  
  aes(x=n, y=scaleSE, col=estimator, pch=estimator) +
  facet_grid( skew ~ df0 , labeller = "label_both" ) +
  geom_line() + geom_point() +
  scale_x_log10( breaks=ns ) + 
  scale_color_viridis_d()
```
We now clearly see the trends.
Overall, the scaled error of the mean is stable across the different distributions. 
The trimmed mean is advantageous when the degrees of freedom are small (note the yellow line is far below the dark line on the left-most facet): we are cropping outliers that destabilize our estimate, which leads to great wins.
As the distribution grows more normal (i.e., as the degrees of freedom increases), this is no longer an advantage and the trimmed mean gets closer to the mean in terms of
performance.
We might think we should be penalized by having dropped 10% of our data, making the standard errors slightly larger: if this were the case, then the effect is not large as the MCSE swamps it (the trimmed mean and mean lines are basically overlapping for df of 15 and larger).

The median, by contrast, is not able to take advantage of the nuances of the individual observations in the data because it is entirely determined by the middle value.
When outliers cause real concern, this cost is minimal.
When outliers are not a concern, the median is just worse.

Overall, we see for precision, the trimmed mean seems an excellent choice: in the presence of outliers it is far more stable than the mean, and when there are no outliers
the cost of using it is small.

Considering the lessons for how we analyze simulation results, our final figure also nicely illustrates how visual displays of simulation results can tell very clear stories.
Eschew complicated tables with lots of numbers.


## Bias-variance tradeoffs

We just looked at the precision of our three estimators, but we did not take into account bias.
In our data generating processes, the median is not the same as the mean.
To see this more clearly, we can generate large datasets for each value of our simulation parameters, then compare the mean and median:

```{r make_big_example_data, cache=TRUE}
#| message: false

source( "case_study_code/trimmed_mean_simulation.R" )

vals <- expand_grid( 
  df0 = unique( results$df0 ),
  skew = unique( results$skew ) 
) %>%
  mutate(
    n = 1e6
  )

res <-
  vals %>%
  mutate(
    data = pmap(., gen.data),
    map_df(data, ~ tibble(mean = mean(.x), median = median(.x), sd = sd(.x)))
  ) %>%
  dplyr::select(-data)

knitr::kable( res, digits=2 )
```

For large skew, the median is not the same as the mean.
If our goal is to estimate the mean, then our median estimator will likewise be systematically off.
The trimmed mean may also suffer: it will trim off extreme right-tail values, but there are no extreme left-tail values to trim, creating imbalance.
However, if the trimmed estimators are much more stable than the mean estimator, we might still prefer to use them. 

Before we just looked at the SE.
But we actually want to know the standard error, bias, and overall root mean squared error (RMSE\index{performance measures!RMSE}).
Can we plot all these on a single plot?  Yes we can!
Following Section @sec-bias-se-rmse-plot, we can make a plot that shows all three measures of estimator performance by pivoting our data longer and then faceting by measure.

```{r}
res2 <- 
  results %>%
  pivot_longer( 
    cols = c( RMSE, bias, SE ), 
    names_to = "Measure",
    values_to = "value" 
  ) %>% 
  mutate(
    Measure = factor( Measure, levels = c("bias","SE","RMSE"))
  )

```


And then we plot, making a facet for each outcome of interest.
We focus on how these trade-offs play out when `df0 = 3`, where we are seeing the largest benefits from trimming in our prior analysis.
We can use lines to show the trends, which is a bit clearer:

```{r}
ns = unique( res2$n )
ggplot( filter( res2, df0 == 3 ), 
        aes(x=n, y=value, col=estimator ) ) +
    facet_grid( skew ~ Measure ) +
    geom_hline( yintercept=0, col="darkgrey" ) +
    geom_line() + geom_point() +
    scale_x_log10( breaks=ns ) +
    labs( y="" ) + 
    scale_color_viridis_d()
```

We could also aggregate without losing too much information.

<!--

We need to bundle because we have too many factors (one now being measure).
We bundle over the four combinations of `df0` by plotting the four individual points for each combination of skew and sample size:

```{r fig.height=4, eval=FALSE}
ns = unique( res2$n )

ggplot( res2 ) +
  aes(x=as.factor(n), y=value, col=estimator ) +
  facet_grid( skew ~ Measure ) +
  geom_hline( yintercept=0, col="darkgrey" ) +
 # geom_boxplot( position = "dodge" ) +
  geom_point(position = position_dodge(width = 0.5)) +
  labs( y="" )
```


-->

<!--

```{r, eval=FALSE}
res2agg <- group_by( res2, n, estimator, skew, Measure ) %>%
  summarise( value = mean(value), .groups = "drop" )

ggplot( filter( res2, df0 == 3 ), aes(x=n, y=value, col=estimator ) ) +
    facet_grid( skew ~ Measure ) +
    geom_hline( yintercept=0, col="darkgrey" ) +
    geom_line() + geom_point() +
    scale_x_log10( breaks=ns ) +
    labs( y="" )
```

-->

The _Bias-SE-RMSE_ plot shows how different estimators trade offdifferent biases and different uncertainties.
The bias is negative for our trimmed estimators because we
are losing the big outliers above and so are getting answers that are too low.

The RMSE captures the overall trade-off in terms of what estimator gives the lowest
overall *error*.
For this distribution, the mean wins as the sample size
increases because the bias basically stays the same and the SE drops.  But
for smaller samples the trimming is superior.
The median (essentially
trimming 50% above and below) is overkill and has too much negative bias.

From a simulation study point of view, notice how we are looking at three
different qualities of our estimators. Some people really care about bias,
some care about RMSE.  By presenting all results we are transparent about how
the different estimators operate.

Next steps would be to also examine the associated estimated standard errors
for the estimators, seeing if these estimates of estimator uncertainty are
good or poor.  This leads to investigation of coverage rates and similar.


## Exercises


### Examine the role of `df0`

How much do your results change when `df0` is larger?  Come up with a visualization that shows the impact, if any, of `df0`.

### Bundle and Aggregate the Bias-SE-RMSE plot

The Bias-SE-RMSE plot above is only for `df0 = 3`.
First make a version that bundles across `df0` values (see Section @sec-bundling).
How many scenarios are in each boxplot?

Now make a version that aggregates across `df0` values (see Section @sec-aggregation).
What is lost, if anything, by this aggregation?



