---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r echo = FALSE, messages = FALSE, include=FALSE}
library(tidyverse)
conflicted::conflicts_prefer(dplyr::select)
conflicted::conflicts_prefer(dplyr::filter)

knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 5,
                      fig.height = 3,
                      out.width = "75%", 
                      fig.align = "center")
options(list(dplyr.summarise.inform = FALSE))
theme_set( theme_classic() )

set.seed( 4404041 )
library(future)
library(furrr)

```

# (PART) Extensions {-}

# Using simulation as a power calculator {#sec:power}

We can use simulation as a power calculator.
In particular, to estimate power, generate data according to a best guess as to what we might find in a planned evaluation, and then analyze these synthetic data and see if we detect the effect we built into our DGP.
We then do this repeatedly, and see how often we detect our effect.
This is power.

Now, if we are generally right about our guesses about our DGP and the associated parameters we plugged into it, in terms of some planned study, then our power will be right on.
This is all a power analysis\index{power analysis} is, using simulation or otherwise.

Simulation has benefits over using power calculators because we can take into account  aspects of data (such as missing values), and also assess non-standard approaches to evaluation that we would be unable to find in a normal power calculator.
We can also compare different proposals for how we would analyze our data, to see which ones are most powerful (and also verify that they are not biased or otherwise broken).

We illustrate these ideas with a case study.
In this example, we are planning a school-level intervention to reduce rates of discipline via an intervention on both teachers and students targeting socio-emotional skills.
For our context, we have strongly predictive historic data and a time-series component.

We are planning a planned randomized controlled trial (RCT),\index{randomized controlled trial} where we will treat entire schools (so this is a cluster-randomized study).
Power is a concern because treating each school is very expensive (we have to run a large training program and also provide coaching of the staff), so treating each unit is a major decision.
Ideally, we want something like 4, 5, or maybe 6 treated schools.
Our diving question is: Can we get away with this?

In our context we know we will be monitoring discipling for two years after treatment.
We also know we will have baseline rates of discipline as well.
After some initial brainstorming, we found a list of different possible regression models we might use, such as using prior years discipline rates as controls, and also using a two-way fixed effects model.
Our second question is then: Are these models valid in this context, and which one is best?



## Getting design parameters from pilot data

Power analyses, via simulation or otherwise, depend on our assessment of what the data will look like when we finally do our analysis.
In our case, we have pilot data from school administrative records (in particular discipline rates for each school and year for a series of five years), and we will use those to estimate parameters to plug into our simulation.
We assume our experimental sample will be on schools that have chronic issues
with discipline, so we filtered our historic data to get those schools most likely be in our study.
Our assumption is that the data we are using for our simulation look like the data we would see (for the control group) when we actually run our experiment.

Here are our data, with each row a potential school in the district:

```{r, messages=FALSE, warnings=FALSE, include=FALSE}
datW = read_csv( "data/discipline_data.csv" )
datW[,2:6] = round( exp( datW[,2:6] ), digits = 3 )
```

```{r}
print( datW, n=4 )
```



## The data generating process

Our first step is to write a data generator that, given a desired number of control and treatment schools, and a hypothesized treatment effect, makes a dataset by sampling vectors of discipline rates, and then imposes a "treatment effect" of scaling the discipline rate by the treatment coefficient for the last two years.

We will generate schools by first estimating a variance-covariance matrix for the 5 years of data on the log-transformed rates (we used the log transform to put things on a multiplicative scale, and to make the distribution of rates more normal given a heavy skew in the original), and then sampling from a multivariate normal distribution with the estimated mean and covariance structure.

We first estimate the mean and covariance structure:
```{r}
datWl = log( datW[ , -1] )
lpd_mns = apply( datWl, 2, mean )
lpd_mns

lpd_cov = cov( datWl )
round( lpd_cov, digits=2 )
```

We then write our DGP function, `make_dat_param()`:

```{r}
make_dat_param = function( n_c, n_t, tx=1 ) {
    n = n_c + n_t
    lpdisc = MASS::mvrnorm( n, mu = lpd_mns, Sigma = lpd_cov )
    lpdisc = exp( lpdisc )
    colnames( lpdisc ) = paste0( "pdisc_", colnames( lpdisc ) )
    lpdisc = dplyr::as_tibble( lpdisc ) %>%
        dplyr::mutate( ID = 1:(dplyr::n()),
                Z = 0 + ( sample( n ) <= n_t ) )
    
    # Add in treatment effect
    lpdisc = dplyr::mutate( lpdisc, 
                     pdisc_2018 = pdisc_2018 * ifelse( Z == 1, tx, 1 ),
                     pdisc_2019 = pdisc_2019 * ifelse( Z == 1, tx, 1 ) )
 
    lpdisc %>% 
      dplyr::relocate( ID, Z )
}
```

Our function generates schools with discipline given by the provided mean and covariance structure; we have calibrated our data generating process to give us data that looks very similar to the data we would see in the field.
For power, realism, in terms of the aspects impacting uncertainty, is key.

For our impact model, the treatment kicks in for the final two years, multiplying discipline rate by `tx` (so `tx = 1` means no treatment effect).

We test our function, hypothesizing the treatment reduces discipline by 50% in the two follow-up years:
```{r}
set.seed( 594585 )
a = make_dat_param( 100, 100, 0.5 )  
print( a, n = 4 )
```

We can group each treatment arm and look at discipline over the years for our dataset:
```{r, echo=FALSE}
datWL <- datW %>% 
  pivot_longer( `2015`:`2019`, names_to = "year",
                values_to = "pdisc" ) %>%
  mutate( year = as.numeric( year ) ) %>%
  group_by( year ) %>%
  summarise( pdisc = mean( pdisc ) )

aL = a %>% 
  pivot_longer( pdisc_2015:pdisc_2019, 
                names_to = c( ".value", "year" ),
                names_pattern = "(.*)_(.*)" ) %>%
  mutate( year = as.numeric( year ) )

aLg = aL %>% group_by( year, Z ) %>%
  summarise( pdisc = mean( pdisc ) )

ggplot( aLg, aes( year, pdisc, col=as.factor(Z) ) ) +
  geom_line() + geom_point() +
  scale_color_manual( values = c("orange","blue", "black" ) ) +
  geom_line( data=datWL, aes( x=year, y=pdisc, col="Real Data" ), lty=2 ) +
  labs( title = "Synthetic discipline data compared to initial data",
        color = "Tx Grp" )
```

Our treatment group drops faster than the control.
We also see the nonlinear structure actually observed in our original data in terms of discipline over time has been replicated (original data is in black).

We next implement all of our candidate treatment effect estimators.
This should all feel very familiar: we are just doing our simulation framework, as usual.

```{r}
eval_dat = function( sdat ) {
    
    # No covariate adjustment, average change model (on log outcome)
    M_raw = lm( log( pdisc_2018 ) ~ 1 + Z, data=sdat )

    # Simple average change model using 2018 as outcome.
    M_simple = lm( pdisc_2018 ~ 1 + Z + pdisc_2017 + pdisc_2016 + pdisc_2015,
                   data=sdat )

    # Simple model on logged outcome
    M_log = lm( log( pdisc_2018 ) ~ 1 + Z + log( pdisc_2017) + 
                  log( pdisc_2016) + log( pdisc_2015 ),
                data=sdat )
    
    # Ratio of average disc to average prior disc as outcome
    sdat = dplyr::mutate( sdat,
                   avg_disc = (pdisc_2018 + pdisc_2019)/2,
                   prior_disc = (pdisc_2017 + pdisc_2016 + pdisc_2015 )/3,
                   disc = pdisc_2018 / prior_disc,
                   disc_two = avg_disc / prior_disc )
    M_ratio = lm( disc ~ 1 + Z, data = sdat )
    M_ratio_twopost = lm( disc_two ~ 1 + Z, data = sdat )
    
    # Use average of two post-tx time periods, averaged to reduce noise
    M_twopost = lm( log( avg_disc ) ~ 1 + Z + log( pdisc_2017 ) + 
                      log( pdisc_2016 ) + log( pdisc_2015 ), 
                    data=sdat )

    # Time and unit fixed effects
    sdatL = tidyr::pivot_longer( sdat, cols = pdisc_2015:pdisc_2019, 
                          names_to = "year",
                          values_to = "pdisc" ) %>%
        dplyr::mutate( Z = Z * (year %in% c( "pdisc_2018", "pdisc_2019" ) ),
                ID = paste0( "S", ID ) )
    
    M_2wfe = lm( log( pdisc ) ~ 0 + ID + year + Z,
                 data=sdatL )

    # Bundle all our models by getting the estimated treatment impact
    # from each.
    models <- list( raw=M_raw, simple=M_simple,
                       log=M_log, ratio = M_ratio, 
                       ratio_twopost = M_ratio_twopost,
                       log_twopost = M_twopost, 
                       FE = M_2wfe )
    rs <- purrr::map_df( models, broom::tidy, .id="model" ) %>%
      dplyr::filter( term=="Z" ) %>%
      dplyr::select( -term ) %>%
      dplyr::arrange( model )
    
    rs
}
```

Our method marches through a host of models; we weren't sure what the gains would be from one model to another, so we decided to conduct power analyses on all of them.
Again, we look at what our evaluation function does:

```{r}
dat = make_dat_param( n_c = 4, n_t = 4, tx = 0.5 )
eval_dat( dat )
```
We have a nice set of estimates, one for each model.  

## Running the simulation

Now we put it all together in our classic simulator:
```{r}
sim_run = simhelpers::bundle_sim( make_dat_param, eval_dat )
```

We then do the usual to run across a set of scenarios, running `sim_run` on each row of the following:
```{r}
res = expand_grid( tx = c( 1, 0.75, 0.5 ),
                   n_c = c( 4, 5, 6, 8, 12, 20 ),
                   n_t = c( 4, 5, 6 ) )
res$reps = 1000
res$seed = 1010203 + 1:nrow(res)
```


```{r secret_run_disc_simulation, include=FALSE}
if ( FALSE ) {
  sim_run( 4, n_c = 20, n_t = 5 )
}
# See below for the presented, visible code
if ( !file.exists( "data/discipline_simulation.rds" ) ) {
  
  library( future )
  library( furrr )
  plan(multisession, workers = parallel::detectCores() - 1 )
  
  res$reps = 10

  fopts = furrr_options(seed = NULL,
                        globals = c("lpd_mns", "lpd_cov") )
  res_u <- simhelpers::evaluate_by_row( params = res,
                               sim_run, 
                               .options = fopts,                 
                               .progress = TRUE )
  
  plan( sequential )
  
  res_u
  
  saveRDS( res_u, file="data/discipline_simulation.rds" )
}
```


For evaluation, we load our saved results and calculate rejection rates (we use an alpha of 0.10 since we are doing one-sided testing):

```{r}
res = readRDS( file="data/discipline_simulation.rds" )

sres <- res %>% group_by( n_c, n_t, tx, model ) %>%
    summarise( E_est = mean( estimate ),
               SE = sd( estimate ),
               E_SE_hat = mean( std.error ),
               pow = mean( p.value <= 0.10 ) ) # one-sided testing
sres
```

## Evaluating power

Once our simulation is run, we can explore power as a function of the design characteristics.
In particular, we eventually want to calculate the chance of noticing effects of different sizes, given various sample sizes we might employ.
Our driving question is how few schools on the treated side can we get away with?
Also, we want to know how much having more schools on the control side allows us to get away with fewer schools on the treated side.


### Checking validity of our models
Before we look at power, we need to check on whether our different models are valid.
This is especiallt important as we are in a small $n$ context, so we know asymptotics may not hold as they should.
To check our models for validity we subset our trials to where `tx = 1`, and look at the rejection rates.

We first run a regression to see if rejection is a function of sample size (are smaller samples more invalid) and treatment-control imbalance.
We center both variables so our intercepts are overall average rejection rates for each model considered:

```{r disc_validity}
sres = mutate( sres,
               n = n_c + n_t,
               imbalance = pmax( n_t / n_c, n_c / n_t ) - 1 )
sres$n = (sres$n - mean(sres$n)) / sd(sres$n)
mod = lm( pow ~ 0 + (n + imbalance) * model - n - imbalance,
          data = filter( sres, tx == 1 ) )
broom::tidy(mod) %>%
  knitr::kable( digits = 3)
```

We can also plot the nominal rejection rates under the null:

```{r}
sres %>% filter( tx == 1 ) %>%
ggplot( aes( n_c, pow, col=model ) ) +
  facet_wrap( ~ n_t, nrow=1 ) +
  geom_line() +
  geom_hline( yintercept = 0.10 ) +
  scale_x_log10(breaks=unique(sres$n_c) )
```

We see the fixed effect models have elevated rates of rejection.
Interestingly, these rates do not seem particularly dependent on sample size or treatment-control imbalance (note lack of significant coefficeints on our regression model).
The other models all appear valid.

We can also check for bias of our methods:
```{r}
sres %>% group_by( model, tx ) %>%
  summarise( E_est = mean( E_est ) ) %>%
  pivot_wider( names_from="tx", values_from="E_est" )
```

We see our models are estimating different things, none of which are the treatment effect as we parameterized it.
In particular, "FE," "log," "raw," and "log_twopost" are all estimating the impact on the log scale.
Note that $log( 0.5 ) \approx -0.69$ and $log( 0.75 ) \approx -0.29$.
Our "simple" estimator is estimating the impact on the absolute scale; reducing discipline rates by 50% corresponds to about a 2% reduction in actual cases.
Finally, "ratio" and "ratio_twopost" are estimating the change in the average ratio of post-policy discipline to pre; they are akin to a gain score as compared to the log regressions.

### Assessing Precision (SE)

Now, which methods are the most precise?
We look at the true standard errors across our methods (we drop "simple" and the "ratio" estimators since they are not on the ratio scale):

```{r disc_precision}
sres %>% 
  group_by( model, n_c, n_t ) %>%
  summarise( SE = mean(SE ) ) %>%
  filter( !(model %in% c( "simple", "ratio", "ratio_twopost" ) ) ) %>%
  ggplot( aes( n_c, SE, col=model )) +
    facet_grid( . ~ n_t ) +
    geom_line() + geom_point() +
    geom_hline( yintercept = 0 ) +
  labs( colour = "Model" )
```

It looks like averaging two years for the outcome is helpful, and bumps up precision.
The two way fixed effects model seems to react to the number of control units differently than the other estimators; it is way more precise when the number of controls is few, but the other estimators catch up.
The "raw" estimator gives a baseline of no covariate adjustment; everything is substantially more precise than it.
The covariates matter a lot.

### Assessing power 

We next look at power over our explored contexts, for the models that we find to be valid (i.e., not FE).  

```{r disc_power}
sres %>% 
  filter( model != "FE",tx != 1 ) %>%
  ggplot( aes( n_c, pow, col=model )) +
    facet_grid(  . ~ tx + n_t, labeller = label_both ) +
    geom_line() + geom_point() +
    geom_hline( yintercept = 0, col="grey" ) +
    geom_hline( yintercept = c( 0.10, 0.80 ), lty=2 ) +
  theme_minimal()+ theme( legend.position="bottom",
                          legend.direction="horizontal",
                          legend.key.width=unit(1,"cm"),
                          panel.border = element_blank() ) +
  labs( title="Power for various methods vs number of controls.",
      y = "Power" )
```

We mark 80% power with a dashed line.  For a 25% reduction in discipline, nothing reaches desired levels of power.
For 50% reduction, some designs do, but we need substantial numbers of control schools.
Averaging two years of outcomes post-treatment seems important: the "twopost" methods have a distinct power bump.
For a single year of outcome data, the log model seems our best bet.


### Assessing Minimum Detectable Effects

Sometimes we want to know, given a design, what size effect we might be able to detect.
The usual measure for this is the Minimum Detectable Effect (MDE), which is usually the size of the smallest effect we could detect with power 80%.

To calculate Minimal Detectable Effects (MDEs) for the log-scale estimators, 
we first average our SEs over our different designs, grouped by sample size, and then convert the SEs to MDEs by multiplying by 2.8.
We then have to convert to our treatment scale by flipping the sign and exponentiating, to get out of the log scale.

```{r disc_mde}
sres2 = sres %>% 
  group_by( model, n_c, n_t ) %>%
  summarise( SE = mean( SE ),
             E_SE_hat = mean( E_SE_hat ) ) %>%
  mutate( MDE = exp( - (1.64 + 0.8) * SE ) )

sres2 %>% 
  filter( !(model %in% c( "simple", "ratio", "ratio_twopost" ) ) ) %>%
  ggplot( aes( n_c, MDE, col=model ) ) +
  facet_wrap( ~ n_t, labeller = label_both ) +
  geom_point() + geom_line()  +
  geom_hline( yintercept = 0.5 ) +
  theme_minimal() +
  scale_x_log10( breaks = unique( sres$n_c ) ) +
  theme( legend.position="bottom",
         legend.direction="horizontal", legend.key.width=unit(1,"cm"),
         panel.border = element_blank() ) +
  labs( x = "Number of control units", y = "MDE (proportion reduction of rate)",
        caption = "A MDE of 0.6 means a 60% reduction (more than half) in discipline rates",
        title = "MDE vs number of control units for various methods" )
```

Corresponding with our findings regarding precision, above, the twopost estimator is the most sensitive, finding the smallest effects.


## Power for Multilevel Data

```{r, include=FALSE}
library( mlmpower )
library( lme4 )

```

Many power analyses are regarding fitting some sort of multilevel model to some type of structured data.
For example, researchers frequently want to calculate power for multisite randomized trials, where each of a series of sites has students randomized to treatment, or not.
Our earlier cluster RCT case study (see, e.g., Chapter \@ref(case-cluster)) is another example of this.

We can use the same simulation framework to calculate power for these types of models.
We write a data generation function that generates our data given our target structure, and then repeatidly generate and analyze data to assess power, just as we have done.

As we saw earlier, however, it can be sometimes tricky to write code that properly has covariates that relate to the different levels of our model, or that divides variance appropriately across levels.
For example, in a multisite experiment, we might want a covariate that has a different mean value within each cluster, but also has variation within cluster.

Instead of immediately writing our own data generation function when faced with such a project, it might be worth looking to the literature to see what tools are available.
In this case, for example, we might come across @enders2023simple, which showcases a package, `mlmpower`, designed to generate data according to a flexible range of multilevel models.

```{r}
library( mlmpower )
model <- outcome('Y') +
  within_predictor('X') +
  between_predictor('W') +
  effect_size(icc = 0.1,
              within = 0.2 )

set.seed(4232345)

dat <- generate( model, 
                 n_within = 5, 
                 n_between = 3 ) %>%
  as_tibble()
dat
cor( dat$X, dat$Y )
```

It even provides methods for calculating power, using a lmer for the estimator (it makes the estimator less visible to the user).
The method in fact runs a multifactor experiment:

```{r mlmpow_calc, echo=FALSE}
pow <- power_analysis(
  model = model,
  replications = 100,
  n_within = c(10,20),
  n_between = c(5,10)
)
pow$power[[1]]
```

We have estimated 93% power for detecting the effect of the within-level predictor, `X`, with 10 students per site and 5 sites, given the parameters we set.
We refer the reader to the package documentation for how to use this specific package.
The broader point is it is sometimes worth digging into provided code to get material for generating data or even running simulations.
That said, each package is designed for a specific purpose, and has its own language.
Here, for example, the effect size defined as `within` is due to the package paying much attention to how covariates are centered and incorporated in the model.
Tying its parameterization to classic regression coefficients may be non-obvious.
It is thus sometimes easy to use a package to get data that looks like data, but does not actually have the structure you intend.
As always, do diagnostics and verification to ensure the tools are working as you expect.

Here, for example, we might generate a very large dataset to get estimated coefficients as a sanity check:

```{r}
dat <- generate( model, 
                 n_within = 100, 
                 n_between = 100 ) %>%
  as_tibble()
M = lme4::lmer( Y ~ X + W + (1|`_id`), data=dat )
arm::display(M)
```

We can check our ICC:
```{r}
sigma2_2 = VarCorr(M)$`_id`[1]
sigma2_1 = attr(VarCorr(M), "sc")^2
sigma2_2 / (sigma2_2 + sigma2_1)
```

Close to the listed 0.10.  More testing is needed.


