# Simulation with clustered data

Generating data with complex structure can be intimidating, but if you set out a recipe for how the data is generated it is often not to bad to build that recipe up with code.

As a simple example, we look at one method for generating clustered data.
This method simulate a cluster sampling frame with the following parameters:

    J = number of clusters in the sampling frame
    n = number of obs per cluster for an even n per J cluster sample
    min.clust.size = minimum simulated cluster size
    max.clust.size = maximum simulated cluster size
    pop.mean = true population mean
    pop.sd = true population standard deviation
    avg.offset = standard deviation of any one cluster's mean from the true population mean

To make things interesting, we will then calculate the mean of the sample in three ways:
     a) sample n subjects from each cluster and calculate an unweighted average
     b) sample n subjects from each cluster and calculate a weighted average (by the actual size of the cluster)
     c) conduct a simple random sample of the entire sampling frame, ignoring clusters 


Also we will calculate the error that each estimate produces two ways:
     a) from the true (parameterized) mean.  This is superpopulation inference.
     b) from the observed (entire sample frame) mean.  This is finite sample inference.

**This is all horrible; need to fix it or remove it**

```{r}
simulation <- function(J = 20, 
                       n = 10, 
                       min.clust.size = 20,
                       max.clust.size = 50,
                       pop.mean = 0, 
                       pop.sd = 1, 
                       avg.offset = 0.1) {
  
  Ca = rnorm( 20 )
  Cb = rnorm( 10000)
  units = c( Ca, Cb )
  
  Sa = sample( Ca, 10 )
  Sb = sample( Cb, 10 )
  
  #Create a vector of cluster sizes of length J by sampling from min to max cluster sizes
  clustsize <- sample(min.clust.size:max.clust.size, J, replace = TRUE)

  #Create a function that creates a vector of observations of the cluster size centered on "pop.mean + offset"  
  #(a randomly generated number to simulate ICC)
    populate.cluster <- function(x) {
      offset <- rnorm(1, mean = 0, sd = avg.offset) #add some noise to create ICC
      rnorm(x, mean = pop.mean + offset, sd = pop.sd)
    }  
  
  #Expand a long list that replicates the cluster ID (1,1,2,2,2,2,3,3,3,3 etc.) by the size of each cluster
  cluster <- as.factor(unlist(mapply(rep, 1:J, clustsize)))

  #Apply the function to create a list of values for each cluster of the appropriate length
  value <- unlist(sapply(clustsize, populate.cluster))
  
  #Merge into a single sampling frame that includes the cluster ID and values for each cluster
  clustersamp <- data.frame(cluster, value)

  #Create a function that subsets a vector by a value (cluster ID) and then draws a sample from that subset
  #and returns the mean of the subset sample and length (number of obs) for the subset
  meanandlength <- function(x){
    subset <- clustersamp$value[clustersamp$cluster == x]
    subset.sample <- sample(subset, n, replace = FALSE)
    mean <- mean(subset.sample)
    length <- length(subset)
    c(mean,length)
  } #Subset the sample to the xth cluster, then take a sample and record the mean
  
  #Apply the meanandlength function across the full sampling frame
  equaln.clustsamp <- data.frame(t(sapply(1:J, meanandlength)))
    colnames(equaln.clustsamp) <- c("mean", "n") #Sample n from each cluster and record mean and length
  
  #Get the weighted mean by the length (size) of each cluster
  weighted.clustsamp.avg <- weighted.mean(equaln.clustsamp$mean, 
                                               equaln.clustsamp$n) 

  #Get the unweighted mean ignoring the difference in cluster size
  unweighted.avg <- mean(equaln.clustsamp$mean) 
  
  #Take a separate simple random sample of size J*n ignoring clusters
  srs.clustsamp <- sample(clustersamp$value, (J*n), replace = FALSE)
  
  #Take the mean of the simple random sample
  srs.clustsamp.avg <- mean(srs.clustsamp)
  
  #Take the mean of the entire cluster sampling frame
  sampframe.mean <- mean(clustersamp$value)
  c(sampframe.mean, srs.clustsamp.avg, weighted.clustsamp.avg, unweighted.avg)
  
}
```

Using our function, we can replicate the formula a bunch of times to generate a dataset of the means by different ways:

```{r}
set.seed(99)
number <- 0
results <- replicate(10000, simulation(J = 30, 
                                       n = 5, 
                                       min.clust.size = 20,
                                       max.clust.size = 50,
                                       pop.mean = 0, 
                                       pop.sd = 1, 
                                       avg.offset = 0))
resultsa <- data.frame ( t(results) )
colnames(resultsa) <- c("sample","simple","weighted","unweighted")

mean(abs(resultsa$simple))
mean(abs(resultsa$weighted))
mean(abs(resultsa$unweighted))

head(resultsa)

error <- data.frame(resultsa$sample - resultsa$simple,
                    resultsa$sample - resultsa$weighted,
                    resultsa$sample - resultsa$unweighted)
colnames(error) <- c("simple","weighted","unweighted")

head(error)

ggplot() +
  geom_density(data = error, aes(simple, col =  "simple")) + 
  geom_density(data = error, aes(weighted, col = "weighted")) +
  geom_density(data = error, aes(unweighted, col = "unweighted"))

mean(abs(error$simple))
mean(abs(error$weighted))
mean(abs(error$unweighted))

t.test(abs(error$simple), abs(error$weighted))

t.test(abs(error$unweighted), abs(error$weighted)) #WEIRD!

sd(error$simple)
sd(error$weighted)
sd(error$unweighted)
```

